---
title: "R Notebook"
output: html_notebook
---

#env setting

```{r,R enviroment activiation,include=FALSE}
renv::activate(project = "~/my_SCP_env")
```

```{r,include=FALSE}
#library(ISCHIA)
library(SCP)
library(ggpubr)
library(future)
library(Seurat)
library(Matrix)
library(dplyr)
library(AnnotationHub)
library(future)
library(future.apply)
library(ggplot2)
library(RColorBrewer)
library(ggsci)
library(gtools)
library(scatterpie)
library(harmony)
library(UCell)
library(tidyr)
library(stringr)
library(dplyr)
library(RColorBrewer)
library(BASS)
library(Lamian)
library(GeneTrajectory)
library(ggVolcano)
library(ggvenn)
library(eulerr)
library(Hmisc)
library(scMetabolism)
library(ggplot2)
library(rsvd)
library(psupertime)
library(ggnewscale)
library(UpSetR)
library(spacexr)
library(tidyverse)
library(clusterProfiler)
library(STdeconvolve)
library(SONAR)
library(EnrichedHeatmap)
library(circlize)
library(CellChat)
library(patchwork)
#library(BioEnricher)
library(SPATAwrappers)
library(SPATA2)
library(scMaSigPro)
```


```{r,multicore setting,include=FALSE}
nbrOfWorkers()
plan(multisession, workers=1)
nbrOfWorkers()
options(Seurat.object.assay.version = 'v4')#### very important!!!!!!
```

# Data preprocessing and zonation annotation

## EDA

```{r,EDA,include=FALSE}
raw.xdbit<- PercentageFeatureSet(raw.xdbit, pattern = "^mt-", col.name = "percent.mt")

#nFeature
raw.xdbit$nGene<-raw.xdbit$nFeature_counts
FeatureStatPlot(raw.xdbit, stat.by = "nGene", group.by = "id", add_box = TRUE,
                palette = "Paired")
ggsave(filename = "R figures/QC/nGene-id.pdf",width = 8,height = 4)


#nCount
raw.xdbit$nCount<-raw.xdbit$nCount_counts

FeatureStatPlot(raw.xdbit, stat.by = "nCount", group.by = "id", add_box = TRUE,
                palette ="Paired")
ggsave(filename = "R figures/QC/nCount-id.pdf",width = 8,height = 4)



# Complexity
raw.xdbit$Complexity<-raw.xdbit$nFeature_counts/raw.xdbit$nCount_counts

FeatureStatPlot(raw.xdbit, stat.by = "Complexity", group.by = "id", add_box = TRUE,
                palette  ="Paired")
ggsave(filename = "R figures/QC/Complexity-id.pdf",width = 8,height = 4)


# percent.mt
FeatureStatPlot(raw.xdbit, stat.by = "percent.mt", group.by = "id", add_box = TRUE,
                palette = "Paired")
ggsave(filename = "R figures/QC/percent mt-id.pdf",width = 8,height = 4)


CellDimPlot(raw.xdbit,group.by = "id",reduction = "umap")

ggsave(filename = "R figures/QC/Dimplot-id before integration.pdf",width = 8,height = 4)

CellDimPlot(raw.xdbit,group.by = "experiment_id",reduction = "umap")

ggsave(filename = "R figures/QC/Dimplot-experiment id before integration.pdf",width = 8,height = 4)

CellDimPlot(raw.xdbit,group.by = "mouse_id",reduction = "umap")

ggsave(filename = "R figures/QC/Dimplot-mouse id before integration.pdf",width = 8,height = 4)
```

## preprocessing

```{r,pp,include=FALSE}
raw.xdbit <- NormalizeData(raw.xdbit,
                          normalization.method = "LogNormalize", 
                          scale.factor = 100000)

raw.xdbit <- FindVariableFeatures(raw.xdbit, 
                                 selection.method = "vst",
                                 nfeatures = 3000)

raw.xdbit<- ScaleData(raw.xdbit,vars.to.regress =c("nGene","nCount","percent.mt"),verbose = F)

raw.xdbit<- RunPCA(raw.xdbit , npcs = 50, verbose = FALSE)

raw.xdbit<- FindNeighbors(object =raw.xdbit  , dims = 1:30, verbose = FALSE)

raw.xdbit <- FindClusters(object = raw.xdbit, reduction.type = "pca", dims.use = 1:30, 
    resolution = 0.1, print.output = FALSE) 
```

## integration

```{r,include=FALSE}
integrated.xDbit<-RunHarmony(raw.xdbit,
                          group.by.vars = "mouse_id",
                           dims.use = 1:30,
                          plot_convergence=T) 

```

```{r,add raw count}
raw.count<-read.csv(file ="raw_counts.csv",header = T )

raw.count<-as.data.frame(t(raw.count))

colnames(raw.count)<-raw.count[1,]

raw.count<-raw.count[-1,]

raw.count <- apply(raw.count, 2, function(x) as.numeric(trimws(x)))

rownames(raw.count)<-rownames(integrated.xDbit)

sparse.count<- as(as.matrix(raw.count), "sparseMatrix")

cell.index<-colnames(integrated.xDbit)

sparse.count<-sparse.count[,which(colnames(sparse.count) %in% cell.index)]

raw.count<-CreateAssayObject(counts = sparse.count, min.cells = 0,
  min.features = 0)

integrated.xDbit[["Raw"]]<-raw.count

save(integrated.xDbit,file = "filter data.RData")
```


```{r,include=FALSE}

integrated.xDbit<- RunUMAP(integrated.xDbit, dims = 1:20,
                    n.neighbors = 100L,
                    min.dist = 0.9,reduction = "harmony")

integrated.xDbit<- FindNeighbors(integrated.xDbit, dims = 1:30,reduction = "harmony")
integrated.xDbit<- FindClusters(
  integrated.xDbit,
  resolution = 0.1,reduction = "harmony"
)

save(integrated.xDbit,file = "integrated xDbit.RData")
```


```{r,vis after integration,include=FALSE}
CellDimPlot(integrated.xDbit,group.by = "id",reduction = "umap")
ggsave(filename = "R figures/Integration and annotation/Dimplot-id after integration.pdf",width = 6,height = 4)

CellDimPlot(integrated.xDbit,group.by = "experiment_id",reduction = "umap", palette = "Dark2")
ggsave(filename = "R figures/Integration and annotation/Dimplot-experiment_id after integration.pdf",width = 6,height = 4)

CellDimPlot(integrated.xDbit,group.by = "mouse_id",reduction = "umap", palette = "Set2")
ggsave(filename = "R figures/Integration and annotation/Dimplot-mouse id after integration.pdf",width = 6,height = 4)

CellDimPlot(integrated.xDbit,group.by = "seurat_clusters",reduction = "umap")
ggsave(filename = "R figures/Integration and annotation/Dimplot-seurat clusters after integration.pdf",width = 6,height = 4)

```

## clustering

```{r,plot spatial plot with seu cluster,include=FALSE}
slice.list<-SplitObject(integrated.xDbit, split.by = "id")
for (i in 1:length(slice.list)) {
  
  tmp<-slice.list[[i]]

CellDimPlot(srt =tmp ,group.by = "seurat_clusters",
              reduction = "spatial",
              bg_color = "black",
              pt.size = 2,
              palette = "Set3",
              theme_use = "theme_blank")+
    ggtitle(names(slice.list[i]))+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(1,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5)) +
    theme(aspect.ratio = 1)

    
ggsave(filename = paste0("R figures/clustering/SpatialClusterPlot-",
                           names(slice.list[i]),
                           ".png"),
       bg = "black", dpi = 150)

ggsave(filename = paste0("R figures/clustering/SpatialClusterPlot-",
                           names(slice.list[i]),
                           ".pdf"), 
       bg = "black",dpi = 150)
  rm(tmp)
  
}

```

## remove low-quality edge spot

```{r,filter,include=FALSE}
integrated.xDbit<-SetIdent(integrated.xDbit,value = "seurat_clusters")
integrated.xDbit<-subset(integrated.xDbit,idents=c( "0","1","2"))
save(integrated.xDbit,file = "filter data.Rdata")
```

```{r,check,include=FALSE}
CellDimPlot(integrated.xDbit,group.by = "seurat_clusters")
ggsave(filename = "R figures/clustering/Dimplot after filtering.pdf", 
      dpi = 150)
ggsave(filename = "R figures/clustering/Dimplot after filtering.png", 
       dpi = 150)
```


## zonation annotation

```{r,check the marker,include=FALSE}
FeatureDimPlot(integrated.xDbit,features = "Cyp2e1",palette = "RdYlBu",
            theme_use = "theme_blank",
            raster = F)#Hep_CV

ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2e1 after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2e1 after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Cyp2f2",palette = "PiYG",
            theme_use = "theme_blank",
            raster = F)#Hep_PV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2f2 after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2f2 after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Igfbp2",palette = "BrBG",
            theme_use = "theme_blank",
            raster = F)#Hep_Mid
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Igfbp2 after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-Igfbp2 after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Spp1",palette = "Greens",
            theme_use = "theme_blank",
            raster = F)#Cholangiocytes
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Spp1 after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Spp1 after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Glul",palette = "RdYlBu",
            theme_use = "theme_blank",
            raster = F)#Hep_CV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Glul after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-Glul after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Lyve1",palette = "BrBG",
            theme_use = "theme_blank",
            raster = F)#Hep_Mid
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Lyve1 after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-Lyve1 after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Krt19",palette = "PiYG",
            theme_use = "theme_blank",
            raster = F)#Hep_PV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Krt19 after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Krt19 after integration.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Hamp",palette = "BrBG",
            theme_use = "theme_blank",
            raster = F)#Hep_PV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Hamp after integration.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Hamp after integration.png",width = 4,height = 4,dpi = 150)
```

```{r,check the marker(harmony),include=FALSE}
FeatureDimPlot(integrated.xDbit,features = "Cyp2e1",palette = "RdYlBu",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Hep_CV

ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2e1 after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2e1 after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Cyp2f2",palette = "PiYG",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Hep_PV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2f2 after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Cyp2f2 after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Igfbp2",palette = "BrBG",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Hep_Mid
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Igfbp2 after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-Igfbp2 after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Spp1",palette = "Greens",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Cholangiocytes
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Spp1 after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Spp1 after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Glul",palette = "RdYlBu",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Hep_CV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Glul after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-Glul after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Lyve1",palette = "BrBG",
            theme_use = "theme_blank",
            raster = F)#Hep_Mid
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Lyve1 after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-Lyve1 after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Krt19",palette = "PiYG",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Hep_PV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Krt19 after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Krt19 after integration(harmony).png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(integrated.xDbit,features = "Hamp",palette = "BrBG",
            theme_use = "theme_blank",reduction = "harmony",
            raster = F)#Hep_PV
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Hamp after integration(harmony).pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot-Hamp after integration(harmony).png",width = 4,height = 4,dpi = 150)
```

```{r,cis ref data,include=FALSE}
 load("H:/Celia group/xDbit(s4) project(renew)/mouse liver altas.RData")


ref.data<-SetIdent(ref.data,value="zonationGroup")
ref.data<-subset(ref.data,idents=c(  "Central", "Mid" ,"Periportal","Portal" ))

ref.data <- NormalizeData(ref.data)
ref.data <- FindVariableFeatures(ref.data)
ref.data <- ScaleData(ref.data)
ref.data <- RunPCA(ref.data)
ref.data <- FindNeighbors(ref.data, dims = 1:30)
ref.data <- FindClusters(ref.data)

CellDimPlot(ref.data,group.by = "zonationGroup")
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for ref data.pdf",width = 4,height = 4,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for ref data.png",width = 4,height = 4,dpi = 150)

FeatureDimPlot(ref.data,features = c("Cyp2f2","Cyp2e1","Glul","Hamp"),ncol = 2)

ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/marker gene Dimplot for ref data.pdf",width = 10,height = 10,dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/marker gene Dimplot for ref data.png",width = 10,height = 10,dpi = 150)
```

### label transfer


```{r,pp for query,include=FALSE}
DefaultAssay(integrated.xDbit)<-"RNA"

db.list<-SplitObject(integrated.xDbit,split.by = "id")
```


```{r,label transfer,include=FALSE}
prediction.list<-list()

for (i in seq_along(db.list)) {
  
  obj<-db.list[[i]]
  
  obj <- NormalizeData(obj)
  obj<- FindVariableFeatures(obj)
  obj <- ScaleData(obj)
  obj <- RunPCA(obj)
  obj <- FindNeighbors(obj, dims = 1:30)
  obj <- FindClusters(obj)
  
  
  db.anchors <- FindTransferAnchors(reference = ref.data, 
                                  query = obj, 
                                  reference.assay ="RNA",
                                  query.assay = "RNA",
                                  dims = 1:30,
                                  reference.reduction = "pca")
  
  
  predictions <- TransferData(anchorset =db.anchors, refdata = ref.data$zonationGroup, dims = 1:30,k.weight =20)
  
  prediction.list[[i]]<-predictions 
  
  
  obj<- AddMetaData(obj, metadata = predictions)
  
  db.list[[i]]<- obj
  
  print(i)
  
  rm(db.anchors,obj, predictions)
  
}



prediction.df<-do.call(rbind,prediction.list)

integrated.xDbit<- AddMetaData(integrated.xDbit, metadata = prediction.df)

integrated.xDbit$predicted.id<-factor(integrated.xDbit$predicted.id,
                                      levels = c("Periportal","Portal" ,
                                                 "Central", "Mid" ))

db.list<-SplitObject(integrated.xDbit,split.by = "id")


pdf(file =  "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/transfer res.pdf",width = 15,height = 5)

for (i in seq_along(db.list)) {
  
 p1<- CellDimPlot(db.list[[i]],group.by = "predicted.id",pt.size = 1,
                    palette = "Set1", reduction = "Spatial")
 
 p2<- CellDimPlot(db.list[[i]],group.by = "predicted.id",pt.size = 1,
                    palette = "Set1", reduction = "harmony")
 
  p3<- CellDimPlot(db.list[[i]],group.by = "predicted.id",pt.size = 1,
                    palette = "Set1", reduction = "umap")
 
 p4<-p1+p2+p3
 
 print(p4)
  
}

dev.off()
```

### fine tune label transfer

```{r, check the marker gene,include=FALSE}
cv.marker<-c("Glul","Lect2","Cyp2e1","Psmd5","Tbx3",
             "Prdx1","Cyp2c55","Pex2","Cyp2c70","Cyp2c37",
             "Cyp2c29","Cyp2c50","Ghr","Gck","Mlxipl",
             "Abcd3","Cyp2c67","Cyp2d9","Fabp1","Cyp2a5",
             "Cyp4a14","Acsl1","Irs1","Aldh4a1",
              "Usp2", "Slco1b2")

names(cv.marker)<-rep("CV.marker",length(cv.marker))

mid.marker<-c("Hamp", "Il1r1", "Cebpb",
              "Thrb", "Ttr", "Apoa1",
             "Scd1", "Ces1f", "Slc38a4")
names(mid.marker)<-rep("Mid.marker",length(mid.marker))

pv.marker<-c("Cyp2f2", "Arg1", "Apoc3","Fga", "Hp",
              "Ass1", "Acly", "Gls2","Atp5k", "Atp5e", "Atp5g3",
             "Mat1a", "Slc25a25","Elovl5", "Acaca",
             "Lpin1", "Adipor2", "Sdhd","Uqcrh", 
             "Idh3b","Pck1", "Insig2")

names(pv.marker)<-rep("PV.marker",length(pv.marker))

marker<-c(cv.marker,mid.marker,pv.marker)

marker.list<-list(PV=pv.marker,
                  Mid=mid.marker,
                  CV=cv.marker)



integrated.xDbit<-AddModuleScore_UCell(integrated.xDbit,features = marker.list,assay = "RNA")

```

```{r,include=FALSE}
integrated.xDbit = FindNeighbors(integrated.xDbit, dims = 1:30,
                          reduction ="harmony")
integrated.xDbit = FindClusters(integrated.xDbit, resolution = 1)

colnames(integrated.xDbit@meta.data)[33]<-"PV.score"
colnames(integrated.xDbit@meta.data)[34]<-"Mid.score"
colnames(integrated.xDbit@meta.data)[35]<-"CV.score"
colnames(integrated.xDbit@meta.data)[35]<-"harmony.snn.1"
```

```{r,vis,include=FALSE}
CellDimPlot(integrated.xDbit,group.by = "harmony.snn.1",label = T,
            reduction = "harmony")
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(harmony).pdf",dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(harmony).png",dpi = 150)

CellDimPlot(integrated.xDbit,group.by = "harmony.snn.1",label = T,
            reduction = "umap")
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(umap).pdf",dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(umap).png",dpi = 150)

CellDimPlot(integrated.xDbit,group.by = "predicted.id",label = T,
            reduction = "harmony")
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for transfer label(harmony).pdf",dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for transfer label(harmony).png",dpi = 150)

CellDimPlot(integrated.xDbit,group.by = "predicted.id",label = T,
            reduction = "umap")
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for transfer label(umap).pdf",dpi = 150)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for transfer label(umap).png",dpi = 150)


ht1 <- GroupHeatmap(integrated.xDbit,
  features = c("PV.score","Mid.score", "CV.score"), 
  fill.by = "group",fill_palette = "Set2",
  group.by = "harmony.snn.1",
  add_violin = TRUE,
  cluster_rows = F,show_column_names = T,
  cluster_columns = T,
  nlabel = 0, show_row_names = TRUE
)$plot

pdf("H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/marker score heatmap for harmony cluster.pdf",width = 10,height = 6)
print(ht1)
dev.off()

ht2 <- GroupHeatmap(integrated.xDbit,
  features =marker, feature_split =  names(marker),
  fill.by = "feature",fill_palette = "Set2",
  group.by = "harmony.snn.1",
  add_violin = TRUE, 
  cluster_rows = F,
  cluster_columns = T,show_column_names = T,
  nlabel = 0, show_row_names = TRUE,
  height=14,width = 7
)$plot


pdf(file =  "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/marker heatmap for harmony cluster.pdf",width = 12,height = 20)
print(ht2)
dev.off()
```

```{r,include=FALSE}
DimPlot(integrated.xDbit,group.by = "harmony.snn.1",
        split.by = "id",ncol = 3,
        cols =brewer.pal(11,"Spectral"),
        label = T,pt.size = 0.5)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(umap) by slice.pdf",dpi = 150,height = 15,width = 8)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(umap) by slice.png",dpi = 150,height = 15,width = 8)

DimPlot(integrated.xDbit,group.by = "harmony.snn.1",
        split.by = "id",ncol = 3,reduction = "harmony",
        cols =brewer.pal(11,"Spectral"),
        label = T,pt.size = 0.5)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(harmony) by slice.pdf",dpi = 150,height = 15,width = 8)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for clusters(harmony) by slice.png",dpi = 150,height = 15,width = 8)
```



```{r,assign the zonation,include=F}
integrated.xDbit<-SetIdent(integrated.xDbit,value = "harmony.snn.1")

integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "0" = "PV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit, 
                              "1" = "Mid")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "2" = "Mid")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "3" = "PV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "4" = "CV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "5" = "Mid")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "6" = "CV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "7" = "PV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "8" = "CV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "9" = "PV")
integrated.xDbit <- RenameIdents(object = integrated.xDbit,
                              "10" = "Mid")


integrated.xDbit$liver.zonation<-integrated.xDbit@active.ident

integrated.xDbit$liver.zonation<-factor(integrated.xDbit$liver.zonation,levels = c("PV","Mid","CV"))

save(integrated.xDbit,file = "filter data.Rdata")
```
```{r,include=FALSE}
CellDimPlot(integrated.xDbit,group.by = "liver.zonation",reduction = "umap", palette = "Set2")

ggsave(filename = "R figures/Integration and annotation/Dimplot-liver zonation after integration(umap).pdf",width =4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-liver zonation after integration(umap).png",width = 4,height = 4,dpi = 150)

CellDimPlot(integrated.xDbit,group.by = "liver.zonation",reduction = "harmony", palette = "Set2")

ggsave(filename = "R figures/Integration and annotation/Dimplot-liver zonation after integration(harmony).pdf",width =4,height = 4,dpi = 150)
ggsave(filename = "R figures/Integration and annotation/Dimplot-liver zonation after integration(harmony).png",width = 4,height = 4,dpi = 150)


ht1 <- FeatureStatPlot(integrated.xDbit,
  stat.by = c("PV.score","Mid.score", "CV.score"),
  legend.position = "top", legend.direction = "horizontal",
  group.by = "liver.zonation", bg.by = "liver.zonation", stack = TRUE,
  palette = "Set2",add_box = T,add_trend = F
)

pdf("H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/marker score heatmap for zonation.pdf",width = 6,height = 8)
print(ht1)
dev.off()

ht2 <-FeatureStatPlot(integrated.xDbit,
  stat.by = marker,
  legend.position = "top", legend.direction = "horizontal",
  group.by = "liver.zonation", bg.by = "liver.zonation", stack = TRUE,
  palette = "Set2",add_box = T
)


pdf(file =  "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/marker heatmap for zonation.pdf",width = 4,height =20)
print(ht2)
dev.off()
```


```{r,include=FALSE}

ht2 <-FeatureStatPlot(integrated.xDbit,
  stat.by = pv.marker,
  legend.position = "top", legend.direction = "horizontal",
  group.by = "liver.zonation", bg.by = "liver.zonation", stack = T,
  palette = "Set2",add_box = T,subtitle = "PV marker"
)



pdf(file =  "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/PV markerheatmap for zonation.pdf",width = 4,height = 25)
print(ht2)
dev.off()

ht2 <-FeatureStatPlot(integrated.xDbit,
  stat.by = cv.marker,
  legend.position = "top", legend.direction = "horizontal",
  group.by = "liver.zonation", bg.by = "liver.zonation", stack = T,
  palette = "Set2",add_box = T,subtitle = "CV marker"
)



pdf(file =  "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/cv markerheatmap for zonation.pdf",width = 4,height = 26)
print(ht2)
dev.off()


ht2 <-FeatureStatPlot(integrated.xDbit,
  stat.by = mid.marker,
  legend.position = "top", legend.direction = "horizontal",
  group.by = "liver.zonation", bg.by = "liver.zonation", stack = T,
  palette = "Set2",add_box = T,subtitle = "Mid marker"
)



pdf(file =  "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/mid markerheatmap for zonation.pdf",width = 4,height =10)
print(ht2)
dev.off()
```


## compare the ground truth and annotated zonation

```{r,include=FALSE}
integrated.xDbit$liver.zonation<-factor(integrated.xDbit$liver.zonation,
                                        levels = c("PV","Mid","CV"))

save(integrated.xDbit,file = "filter data.Rdata")

slice.list<-SplitObject(integrated.xDbit, split.by = "id")
```

```{r,Gene Spatial plot,include=FALSE}

col<-brewer.pal(12,"Paired")
col<-col[c(1,11,5)]
for (i in 1:length(slice.list)) {
  
  tmp<-slice.list[[i]]
  
  CellDimPlot(srt =tmp ,group.by = "liver.zonation",
              reduction = "spatial",
              bg_color = "black",
              pt.size = 2.5,
             palcolor = col,
              theme_use = "theme_blank")+
    ggtitle(names(slice.list[i]))+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(0.5,0.5,0.5,0.5), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5))
  
  ggsave(filename = paste0("R figures/zonation validation/SpatialPlot-",
                           names(slice.list[i]),
                           ".pdf"),dpi = 150)
  
  ggsave(filename = paste0("R figures/zonation validation/SpatialPlot-",
                           names(slice.list[i]),
                           ".png"),dpi = 150)
  rm(tmp)
  
}

for (i in 1:length(slice.list)) {
  
  tmp<-slice.list[[i]]

col<-brewer.pal(12,"Paired")
col<-col[c(1,11,5)]
p1<-CellDimPlot(srt =tmp ,group.by = "liver.zonation",
              reduction = "spatial",
              bg_color = "black",
              pt.size = 2.5,
             palcolor = col,
              theme_use = "theme_blank")+
    ggtitle(names(slice.list[i]))+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(1,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5)) +
    theme(aspect.ratio = 1)


 p2<- FeatureDimPlot(tmp,features = "Cyp2e1",palette = "viridis",
                     slot = "data",
                  
                     lower_quantile = 0,
  upper_quantile = 0.75,
                    
            theme_use = "theme_blank",reduction = "spatial",
            pt.size = 2.5,
            raster = F)+
    ggtitle("Cyp2e1")+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(1,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5,size = 12))+
    theme(aspect.ratio = 1)
 
  p3<- FeatureDimPlot(tmp,features = "Hamp",palette = "viridis",
                       slot = "data",
                     
                      lower_quantile = 0,
  upper_quantile = 0.75,
                   
                     
            theme_use = "theme_blank",reduction = "spatial",
            pt.size = 2.5,
            raster = F)+
    ggtitle("Hamp")+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(1,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5,size = 12))+
    theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"))
  
    p4<- FeatureDimPlot(tmp,features = "Cyp2f2",palette = "viridis",
            theme_use = "theme_blank",reduction = "spatial",
             slot = "data",
           
            lower_quantile = 0,
  upper_quantile = 0.75,
            pt.size = 2.5,
            raster = F)+
    ggtitle("Cyp2f2")+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(0.2,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5,size = 12))+
    theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"))


    
ggsave(filename = paste0("R figures/zonation validation/SpatialGenePlot-",
                           names(slice.list[i]),
                           ".png"), 
       ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2), 
       bg = "black", width = 12, height = 12,dpi = 150)

ggsave(filename = paste0("R figures/zonation validation/SpatialGenePlot-",
                           names(slice.list[i]),
                           ".pdf"), 
       ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2), 
       bg = "black", width = 12, height = 12,dpi = 150)
  rm(tmp)
  
}

col<-brewer.pal(12,"Paired")
col<-col[c(1,11,5)]
DimPlot(integrated.xDbit,group.by = "liver.zonation",
        split.by = "id",ncol = 3,
        cols =col,
        label = T,pt.size = 0.5)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for zonation(umap) by slice.pdf",dpi = 150,height = 15,width = 8)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for zonation(umap) by slice.png",dpi = 150,height = 15,width = 8)

DimPlot(integrated.xDbit,group.by = "liver.zonation",
        split.by = "id",ncol = 3,reduction = "harmony",
        cols =col,
        label = T,pt.size = 0.5)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for zonation(harmony) by slice.pdf",dpi = 150,height = 15,width = 8)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/Integration and annotation/Dimplot for zonation(harmony) by slice.png",dpi = 150,height = 15,width = 8)
```

```{r,include=FALSE}
pdf(file ="R figures/zonation validation/Spatial nCount per slice.pdf", 
    width = 12, height = 6)
for (i in 1:length(slice.list)) {
  
  tmp<-slice.list[[i]]

col<-brewer.pal(12,"Paired")
col<-col[c(1,11,5)]
p1<-CellDimPlot(srt =tmp ,group.by = "liver.zonation",
              reduction = "spatial",
              bg_color = "black",
              pt.size = 2.5,
             palcolor = col,
              theme_use = "theme_blank")+
    ggtitle(names(slice.list[i]))+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(1,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5)) +
    theme(aspect.ratio = 1)


 p2<- FeatureDimPlot(tmp,features = "nCount",palette = "viridis",
            theme_use = "theme_blank",reduction = "spatial",
            pt.size = 2.5,
            raster = F)+
    ggtitle("nCount")+
    theme(plot.background = element_rect(fill = "black", color = "black"),
          panel.background = element_rect(fill = "black", color = "black"),
          plot.margin = unit(c(1,1,1,1), "lines"), 
          text = element_text(color = "white"), 
          axis.title = element_text(color = "white"), 
          axis.text = element_text(color = "white"),
          legend.title = element_text(color = "white"), 
          legend.text = element_text(color = "white"), 
          plot.title = element_text(color = "white", hjust = 0.5,size = 12))+
    theme(aspect.ratio = 1)
 

    
ggsave(filename = paste0("R figures/zonation validation/SpatialGenePlot-nCount-",
                           names(slice.list[i]),
                           ".png"), 
       ggarrange(p1, p2, ncol = 2, nrow = 1), 
       bg = "black", width = 12, height = 6,dpi = 150)

ggsave(filename = paste0("R figures/zonation validation/SpatialGenePlot-",
                           names(slice.list[i]),
                           ".pdf"), 
       ggarrange(p1, p2,  ncol = 2, nrow = 1), 
       bg = "black", width = 12, height = 6,dpi = 150)
p<-p1+p2

print(p)

  rm(tmp)
  
}
dev.off()
```




## DEG

```{r,include=FALSE}
#DEG per zonation

integrated.xDbit <- RunDEtest(srt =integrated.xDbit, group_by ="liver.zonation", fc.threshold = 1, only.pos = FALSE,test.use = "MAST",latent.vars = "mouse_id")

# pathway enrichment
integrated.xDbit<- RunEnrichment(
  srt = integrated.xDbit, group_by = "liver.zonation", db = "GO_BP", species = "Mus_musculus",test.use = "MAST",GO_simplify = T,
  DE_threshold = "avg_log2FC > 0.5 & p_val_adj < 0.05"
)

# GSEA
integrated.xDbit <- RunGSEA(
  srt = integrated.xDbit, group_by = "liver.zonation", db = "GO_BP", species = "Mus_musculus",
  DE_threshold =  "avg_log2FC > 0.5 & p_val_adj < 0.05",test.use = "MAST",GO_simplify = T,
)

integrated.xDbit<- AnnotateFeatures(integrated.xDbit, species = "Mus_musculus", db = c("TF", "CSPA"))

save(integrated.xDbit,file = "filter data.Rdata")
```

```{r,include=FALSE}
DEGs <- integrated.xDbit@tools$DEtest_liver.zonation$AllMarkers_MAST
DEGs <- DEGs[with(DEGs, avg_log2FC > 1 & p_val_adj < 0.05), ]


ht <- GroupHeatmap(
  srt = integrated.xDbit, 
  group.by = "liver.zonation", features = DEGs$gene, feature_split = DEGs$group1,
  feature_annotation = c("TF", "CSPA"),
  feature_annotation_palcolor = list(c("gold", "steelblue"), c("forestgreen")),
  heatmap_palette = "RdYlBu",nlabel = 60,lib_normalize = T,libsize = 1,
  height = 8, width = 4
)

pdf(file ="R figures/Integration and annotation/heatmap for zonation(group).pdf",height = 10,width = 10 )

print(ht$plot)

dev.off()

ht <- FeatureHeatmap(
  srt = integrated.xDbit, 
  group.by = "liver.zonation", features = DEGs$gene, feature_split = DEGs$group1,
  feature_annotation = c("TF", "CSPA"),
  feature_annotation_palcolor = list(c("gold", "steelblue"), c("forestgreen")),
  heatmap_palette = "RdYlBu",nlabel = 60,lib_normalize = T,libsize = 1,
  height = 8, width = 4
)

pdf(file ="R figures/Integration and annotation/heatmap for zonation(feature).pdf",height = 10,width = 10 )

print(ht$plot)

dev.off()


ht <- GroupHeatmap(
  srt = integrated.xDbit, 
  group.by = "liver.zonation", features = DEGs$gene, feature_split = DEGs$group1,
  feature_annotation = c("TF", "CSPA"),split.by = "age",
  feature_annotation_palcolor = list(c("gold", "steelblue"), c("forestgreen")),
  heatmap_palette = "RdYlBu",nlabel = 60,lib_normalize = T,libsize = 1,
  height = 8, width = 4
)

pdf(file ="R figures/Integration and annotation/heatmap for zonation(group)  splited by aging.pdf",height = 10,width = 10 )

print(ht$plot)

dev.off()

ht <- FeatureHeatmap(
  srt = integrated.xDbit, 
  group.by = "liver.zonation", features = DEGs$gene, feature_split = DEGs$group1,
  feature_annotation = c("TF", "CSPA"),split.by = "age",
  feature_annotation_palcolor = list(c("gold", "steelblue"), c("forestgreen")),
  heatmap_palette = "RdYlBu",nlabel = 60,lib_normalize = T,libsize = 1,
  height = 8, width = 4
)

pdf(file ="R figures/Integration and annotation/heatmap for zonation(feature) splited by aging.pdf",height = 10,width = 10 )

print(ht$plot)

dev.off()

```

```{r,include=FALSE}
pdf(file = "R figures/Integration and annotation/Enrichemnt plot (PV).pdf",width = 10,height = 8)
EnrichmentPlot(
  srt =integrated.xDbit, group_by ="liver.zonation", group_use = "PV",
  plot_type = "network",test.use = "MAST"
)
dev.off()

pdf(file = "R figures/Integration and annotation/Enrichemnt plot (CV).pdf",width = 10,height = 8)
EnrichmentPlot(
  srt =integrated.xDbit, group_by ="liver.zonation", group_use = "CV",
  plot_type = "network",test.use = "MAST"
)
dev.off()

pdf(file = "R figures/Integration and annotation/Enrichemnt plot (Mid).pdf",width = 10,height = 8)
EnrichmentPlot(
  srt =integrated.xDbit, group_by ="liver.zonation", group_use = "Mid",
  plot_type = "network",test.use = "MAST"
)
dev.off()

pdf(file = "R figures/Integration and annotation/Enrichemnt plot (comparison).pdf",width = 10,height = 8)
EnrichmentPlot(
  srt =integrated.xDbit, group_by ="liver.zonation", 
  plot_type = "comparison",test.use = "MAST"
)
dev.off()
```

# Construct PV-CV axis(Trajectory) 

```{r,load the data}
load("H:/Celia group/xDbit project/filter data.RData")

DefaultAssay(integrated.xDbit)<-"RNA"

slice.list<-SplitObject(integrated.xDbit, split.by = "id")
```

```{r,write a normalization function}
normalize.trajectory <- function(x) {
  # Normalize x to [0, 1]
  normalized_x = (x - min(x)) / (max(x) - min(x))
  
  # Scale to [-1, 1]
  scaled_x = normalized_x * 2 - 1
  
  return(scaled_x)
}
```


```{r,run psupertime,include=FALSE}
trajectory.order.list<-list()

psuper.obj.list<-list()

set.seed(53144)

for (i in 1:length(slice.list)) {


tmp<-slice.list[[i]]

# normalize the data

DefaultAssay(tmp)<-"RNA"

tmp <- NormalizeData(tmp, normalization.method = "LogNormalize", 
                          scale.factor = 100000)

tmp<- FindVariableFeatures(tmp,selection.method = "vst",
                              nfeatures = 2000)

tmp<- ScaleData(tmp,vars.to.regress =c("nGene","nCount","percent.mt"),verbose = F)

tmp<-as.SingleCellExperiment(tmp)

# Run psupertime

y=factor(tmp$liver.zonation,levels=c("PV","Mid","CV"))

set.seed(53144)

psuper_obj  = psupertime(tmp, y,n_folds = 10,min_expression =0.1,
                         sel_genes = "all",
                         score = "xentropy",test_propn=0.3)

psuper_obj$proj_dt$psuper<-normalize.trajectory(psuper_obj$proj_dt$psuper)
psuper_obj$cuts_dt[,1]<-NA

plot_labels_over_psupertime(psuper_obj, 
                            label_name='liver.zonation',
                            palette = "Set2")+
  xlab("PV-CV Trajectory")+xlim(-1.5,1.5)+
  ggtitle(tmp$id)+
  theme(plot.title = element_text(hjust = 0.5, 
                                  vjust = 1,
                                  size = 20, 
                                  face = "bold"),
    axis.title.x = element_text(face = "bold", size = 15), 
    axis.title.y = element_text(face = "bold", size = 15))

ggsave(filename =paste0("R figures/psupertime/Density plot-",unique(tmp$id),".pdf"),dpi = 150,height = 4,width = 6)

ggsave(filename =paste0("R figures/psupertime/Density plot-",unique(tmp$id),".png"),dpi = 150,height = 4,width = 6)

plot_identified_gene_coefficients(psuper_obj)+
  ggtitle(tmp$id)+xlab("")+
  theme(plot.title = element_text(hjust = 0.5, 
                                  vjust = 1, 
                                  size = 20, 
                                  face = "bold"),
        axis.title.y = element_text(face = "bold", size = 15))

ggsave(filename =paste0("R figures/psupertime/Gene Coefficience-",unique(tmp$id),".pdf"),dpi = 150,height = 4,width = 6)

ggsave(filename =paste0("R figures/psupertime/Gene Coefficience-",unique(tmp$id),".png"),dpi = 150,height = 4,width = 6)


plot_identified_genes_over_psupertime(psuper_obj, 
                                      label_name='liver.zonation',
                                      palette = "Set2")+
  ylab("PV-CV Trajectory")+
  ggtitle(tmp$id)+
  theme(plot.title = element_text(hjust = 0.5, 
                                  vjust = 1, 
                                  size = 20, 
                                  face = "bold"),
    axis.title.x = element_text(face = "bold", size = 15), 
    axis.title.y = element_text(face = "bold", size = 15))

ggsave(filename =paste0("R figures/psupertime/Feature Trajectory plot-",unique(tmp$id),".pdf"),dpi = 150,height = 6,width = 8)

ggsave(filename =paste0("R figures/psupertime/Feature Trajectory plot-",unique(tmp$id),".png"),dpi = 150,height = 6,width = 8)


trajectory.order<-psuper_obj$proj_dt[,c(1:2)]

trajectory.order.list[[i]]<-trajectory.order
names(trajectory.order.list)[i]<-as.character(unique(tmp$id))


psuper.obj.list[[i]]<-psuper_obj
names(psuper.obj.list)[i]<-as.character(unique(tmp$id))

rm(tmp,psuper_obj,trajectory.order,y)

gc()

}

save(trajectory.order.list,psuper.obj.list,file = "psupertime output.RData")

```

```{r,add metadata to seurat,include=FALSE}
trajectory.mat<-as.data.frame(do.call(rbind,trajectory.order.list))

rownames(trajectory.mat)<-trajectory.mat$cell_id

colnames(trajectory.mat)[2]<-"PV.CV.Trajectory"

integrated.xDbit<-AddMetaData(integrated.xDbit,trajectory.mat)

save(integrated.xDbit,file = "filter data.RData")

```

```{r,Umap plot for all slice,include=FALSE}
#slice.list<-SplitObject(integrated.xDbit, split.by = "id")

integrated.xDbit$PV.CV.Trajectory<-as.numeric(integrated.xDbit$PV.CV.Trajectory)

FeatureDimPlot(integrated.xDbit,features ="PV.CV.Trajectory",
               palette = "rocket",bg_cutoff = -1,
               theme_use ="theme_blank",
            reduction ="umap" )+
      ggtitle(c("PV-CV axis"))+
      theme(plot.background = element_rect(fill = "black", color = "black"),
            panel.background = element_rect(fill = "black", color = "black"),
            plot.margin = unit(c(1,1,1,1), "lines"), 
            text = element_text(color = "white"), 
            axis.title = element_text(color = "white"), 
            axis.text = element_text(color = "white"),
            legend.title = element_text(color = "white"), 
            legend.text = element_text(color = "white"), 
            plot.title = element_text(color = "white", hjust = 0.5,
                                      face = "bold",size = 20))+
      theme(aspect.ratio = 1) 

ggsave(filename = "R figures/psupertime/Trajectory plot(umap).pdf",width = 5,height = 4)
ggsave(filename = "R figures/psupertime/Trajectory plot(umap).png",width = 5,height = 4)

FeatureDimPlot(integrated.xDbit,features ="Cyp2f2",
               palette = "Spectral",
               theme_use ="theme_blank",
            reduction ="umap" )+
      ggtitle(c("Cyp2f2"))+
      theme(plot.background = element_rect(fill = "black", color = "black"),
            panel.background = element_rect(fill = "black", color = "black"),
            plot.margin = unit(c(1,1,1,1), "lines"), 
            text = element_text(color = "white"), 
            axis.title = element_text(color = "white"), 
            axis.text = element_text(color = "white"),
            legend.title = element_text(color = "white"), 
            legend.text = element_text(color = "white"), 
            plot.title = element_text(color = "white", hjust = 0.5,
                                      face = "bold",size = 20))+
      theme(aspect.ratio = 1) 

ggsave(filename = "R figures/psupertime/Trajectory plot-Cyp2f2(umap).pdf",width = 5,height = 4)
ggsave(filename = "R figures/psupertime/Trajectory plot-Cyp2f2(umap).png",width = 5,height = 4)

FeatureDimPlot(integrated.xDbit,features ="Cyp2e1",
               palette = "Spectral",
               theme_use ="theme_blank",
            reduction ="umap" )+
      ggtitle(c("Cyp2e1"))+
      theme(plot.background = element_rect(fill = "black", color = "black"),
            panel.background = element_rect(fill = "black", color = "black"),
            plot.margin = unit(c(1,1,1,1), "lines"), 
            text = element_text(color = "white"), 
            axis.title = element_text(color = "white"), 
            axis.text = element_text(color = "white"),
            legend.title = element_text(color = "white"), 
            legend.text = element_text(color = "white"), 
            plot.title = element_text(color = "white", hjust = 0.5,
                                      face = "bold",size = 20))+
      theme(aspect.ratio = 1) 

ggsave(filename = "R figures/psupertime/Trajectory plot-Cyp2e1(umap).pdf",width = 5,height = 4)
ggsave(filename = "R figures/psupertime/Trajectory plot-Cyp2e1(umap).png",width = 5,height = 4)

CellDimPlot(integrated.xDbit,group.by = "liver.zonation",
               palette = "Set2",
               theme_use ="theme_blank",
            reduction ="umap" )+
      ggtitle(c("Liver Zonation"))+
      theme(plot.background = element_rect(fill = "black", color = "black"),
            panel.background = element_rect(fill = "black", color = "black"),
            plot.margin = unit(c(1,1,1,1), "lines"), 
            text = element_text(color = "white"), 
            axis.title = element_text(color = "white"), 
            axis.text = element_text(color = "white"),
            legend.title = element_text(color = "white"), 
            legend.text = element_text(color = "white"), 
            plot.title = element_text(color = "white", hjust = 0.5,
                                      face = "bold",size = 20))+
      theme(aspect.ratio = 1) 

ggsave(filename = "R figures/psupertime/Trajectory plot-liver.zonation(umap).pdf",width = 5,height = 4)
ggsave(filename = "R figures/psupertime/Trajectory plot-liver.zonation(umap).png",width = 5,height = 4)
```

```{r,SpatialTrajectoryPlot for each slide,include=FALSE}
source("SpatialTrajectoryPlot.R")

slice.list<-slice.list[-15]
pdf(file = "R figures/psupertime/Combined Trajectory plot.pdf",bg = "black",width = 22,height =20 )
SpatialTrajectoryPlot(srt.list = slice.list,feature = "PV.CV.Trajectory",
                     assay ="RNA",
                     palette = "rocket" )
dev.off()

for (i in 1:length(selected.feature)) {
  gene<-selected.feature[i]
  
  p<-SpatialGenePlot(srt.list = slice.list,feature = gene)
  
  ggsave(filename = paste0("R figures/CV/SpatialGenePlot-",gene,".png"), p,
       limitsize = FALSE, bg = "black", width = 12, height = 50,dpi = 150)


  ggsave(filename =  paste0("R figures/CV/SpatialGenePlot-",gene,".pdf"), p,
       limitsize = FALSE, bg = "black", width = 12, height = 50,dpi = 150)
  
}

```

```{r,Denoise for the plot,include=FALSE}
source("CreateDenoisedAssay.R")

integrated.xDbit<-CreateDenoisedAssay(srt = integrated.xDbit,slot = "counts",new.assay.name = "DeNoise")

save(integrated.xDbit,file = "filter data.RData")
```


```{r,zonation specific gene,include=FALSE}
source("DbDynamicsPlot.R")

top.feature<-c("Cyp2a5","Serpina6","Cyp2f2","Cyp2e1","Mup7","Gulo","Spp1",
               "Elovl3","Gstm2","Hsd3b5","Ugt2b1","Acaa1b")

for (i in 1:length(top.feature)) {
  
  DbDynamicsPlot(srt = integrated.xDbit,feature = top.feature[i],assay = "DeNoise")
  
  ggsave(filename = paste0("R figures/zonation specific gene by psupertime/Zonation gene-",top.feature[i],".pdf"),
         width = 5,height = 4,dpi = 150)
  ggsave(filename = paste0("R figures/zonation specific gene by psupertime/Zonation gene-",top.feature[i],".png"),
          width = 5,height = 4,dpi = 150)
}

```

# scMagSigPro at RNA
## run scMagSigPro
```{r,include=FALSE}
DefaultAssay(integrated.xDbit)<-"RNA"
xdbit.sce<-as.SingleCellExperiment(integrated.xDbit)
```

```{r,include=FALSE}
xdbit.mag.obj<-as_scmp(object = xdbit.sce,from = "sce",
                       align_pseudotime = T,
                       verbose = T,
                       additional_params = list(labels_exist = TRUE,
                                                exist_ptime_col = "PV.CV.Trajectory",
                                                exist_path_col = "age"))

xdbit.mag.obj <- sc.squeeze(xdbit.mag.obj)

rm(xdbit.sce)
```

```{r,plot bin,include=FALSE}
p<-plotBinTile(xdbit.mag.obj)+
  ggtitle("Binned Pseudotime by scMagSigPro")+
  ylab("Group")+xlab("Binned Pseudotime ")+
  scale_fill_viridis(name = "Bin Size")+
  theme(plot.title = element_text(face = "bold",size = 14,hjust = 0.5),
        axis.title =element_text(face = "bold",size = 12),
        axis.text = element_text(size = 12))
ggsave(p,filename = "R figures/scMagSigPro/pseudotime bins.pdf",width = 6,height = 4,dpi = 300)
ggsave(p,filename = "R figures/scMagSigPro/pseudotime bins.png",width = 6,height = 4,dpi = 300)  
```

```{r,set degree,include=FALSE}
xdbit.mag.obj<- sc.set.poly(xdbit.mag.obj, poly_degree =3)

save(xdbit.mag.obj,file ="scMagSigPro obj.RData" )
```

```{r,run scMagSigPro,include=FALSE}
# Detect non-flat profiles
xdbit.mag.obj <- sc.p.vector(xdbit.mag.obj,
                             offset = TRUE, 
                             p_value = 0.05, 
                             min_na = 10,
                             verbose =T,
                             log_offset = TRUE
)

# Model refinement

xdbit.mag.obj <- sc.t.fit(xdbit.mag.obj, verbose = TRUE)

save(xdbit.mag.obj,file ="scMagSigPro obj.RData" )
```

```{r,get sig gene,include=FALSE}
xdbit.mag.obj <- sc.filter(
  scmpObj = xdbit.mag.obj ,
  rsq = 0.7,
  vars = "groups",
  intercept = "dummy",
  includeInflu = F
)
```


```{r,cluster the sig scMagSigPro gene,include=FALSE}
xdbit.mag.obj<- sc.cluster.trend(
  xdbit.mag.obj,
  cluster_by = "counts",
  geneSet = "union",
  k =6
)

p<-plotTrendCluster(
  scmpObj =xdbit.mag.obj,
  plot = "counts",
  logs = TRUE,
  summary_mode='mean',
  smoothness = 0.01,
  verbose = FALSE,
  lines = TRUE
)

p<-p+ggtitle("Dynamical Gene Expression over PV-CV axis")+ylab("Binned Pseudobulk Expression")+
  xlab("Pooled PV-CV Trajectory")+
  theme(plot.title = element_text(face = "bold",size = 14,hjust = 0.5),
        axis.text.x = element_text(angle = 0),
        axis.title =element_text(face = "bold",size = 12),
        axis.text = element_text(size = 12))


ggsave(p,filename = "R figures/scMagSigPro/clustered scMagSigPro Sig Genes.png",width = 7,height = 5,dpi = 300) 
ggsave(p,filename = "R figures/scMagSigPro/clustered scMagSigPro Sig Genes.pdf",width = 7,height = 5,dpi = 300) 
```

```{r,extract the sig genes}
sig.mag.gene<-as.data.frame( do.call(rbind,xdbit.mag.obj@Significant@clusters))

sig.mag.gene$gene=rownames(sig.mag.gene)
colnames(sig.mag.gene)[1]<-"cluster"

save(sig.mag.gene,file = "sig scMagSigPro gene.RData")
write.csv(sig.mag.gene,file = "sig scMagSigPro gene.csv")
```

```{r,add lineage for seu}
integrated.xDbit$Young.Trajectory<-integrated.xDbit$PV.CV.Trajectory
integrated.xDbit$Young.Trajectory[which(integrated.xDbit$age=="old")]<-NA

integrated.xDbit$Old.Trajectory<-integrated.xDbit$PV.CV.Trajectory
integrated.xDbit$Old.Trajectory[which(integrated.xDbit$age=="young")]<-NA
```

```{r,include=FALSE}
integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =sig.mag.gene$gene,assay = "DeNoise",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))
```

```{r}
cluster.gene<-paste0("Cluster_",sig.mag.gene$cluster)
names(cluster.gene)<-sig.mag.gene$gene

integrated.xDbit <- AnnotateFeatures(integrated.xDbit, species = "Mus_musculus", db = c("TF", "CSPA"))

ht <- DynamicHeatmap(
  srt =integrated.xDbit, 
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = 25, 
  feature_split = cluster.gene,
  cluster_rows = TRUE,
  min_expcells = 1,
  r.sq = 0,
  dev.expl = 0,
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  species = "Mus_musculus", db = c("GO_BP","Reactome"),
  anno_terms = TRUE, 
  anno_keys = F, 
  anno_features = F,
  feature_annotation = c("TF", "CSPA"),
  feature_annotation_palcolor = list(c("gold", "steelblue"), c("forestgreen")),
  topTerm =6,
  height = 8, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/dynamics heatmap.pdf",width = 16,height = 18)
print(ht$plot)
dev.off()

save(integrated.xDbit,file = "filter data.Rdata")

save(ht,file="heatmap obj for scMagSigPro.RData")
```
## TCA

```{r,include=FALSE}
# Extract TCA associated gene and calcuate the module score
load("heatmap obj for scMagSigPro.RData")

enrich.res<-ht$enrichment$enrichment

enrich.res<-enrich.res[which(enrich.res$Database=="Reactome" &
                               enrich.res$p.adjust<0.05),]

tca.res<-enrich.res[which(enrich.res$Description=="The citric acid (TCA) cycle and respiratory electron transport"),]

tca.gene<-str_split(tca.res$geneID,"/")[[1]]

tca.gene.list<-list("TCA.Module.Score"=tca.gene)

integrated.xDbit<-AddModuleScore_UCell(integrated.xDbit,features = tca.gene.list,BPPARAM = BiocParallel::SnowParam(workers = 1))

# dynamic heatmap for all associated gene

integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =tca.gene,
                                       assay = "DeNoise",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))

ht1 <- DynamicHeatmap(
  srt =integrated.xDbit, features = tca.gene,
  assay = "DeNoise",
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = 25, 
  cluster_rows = TRUE,
  min_expcells = 1,
  nlabel = 10,
  show_row_names = F,
  r.sq = 0,
  dev.expl = 0,
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  topTerm =6,
  height = 4, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/TCA/dynamics heatmap for TCA.pdf",width = 8,height = 6)
print(ht1$plot)
dev.off()
```



```{r,lm for ploidy,include=FALSE}
source("ScatterHistoPlot.R")
p<-ScatterHistoPlot(seu = integrated.xDbit,
                 var = "TCA.Module.Score_UCell",
                 plot.x = "nuclei.2n",
                 xlab = "2n Probability",
                 plot.y = "TCA.Module.Score_UCell",
                 ylab = "TCA Module Score")

ggsave(p,filename = "R figures/scMagSigPro/TCA/lm between 2n and TCA score.pdf",height = 4,width = 6)
ggsave(p,filename = "R figures/scMagSigPro/TCA/lm between 2n and TCA score.png",height = 4,width = 6)

p<-ScatterHistoPlot(seu = integrated.xDbit,
                 var = "TCA.Module.Score_UCell",
                 plot.x = "nuclei.4n",
                 xlab = "4n Probability",
                 plot.y = "TCA.Module.Score_UCell",
                 ylab = "TCA Module Score")

ggsave(p,filename = "R figures/scMagSigPro/TCA/lm between 4n and TCA score.pdf",height = 4,width = 6)
ggsave(p,filename = "R figures/scMagSigPro/TCA/lm between 4n and TCA score.png",height = 4,width = 6)
```




```{r,TCA asso genes for cluster 2,include=FALSE}
integrated.xDbit$age<-factor(integrated.xDbit$age,
                             levels = c("young","old"))

p<-FeatureStatPlot(integrated.xDbit,group.by = "liver.zonation",
                split.by = "age",palette ="lancet",
                stat.by = c("Cox5a",
                            "Ndufs1",
                            "Rxra",
                            "Fh1",
                            "Pdha1",
                            "Atp5d"),
                assay = "DeNoise",
                add_box = T,stack = T
                )+
  ggtitle("Mitochondria Dysfunction")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/TCA/TCA gene.pdf",
       height = 10,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/TCA/TCA gene.png",
       height = 10,width = 6,dpi = 300)


source("DbDynamicsPlot.R") 

gene.idx<-c("Cox5a","Ndufs1","Rxra","Fh1", "Pdha1","Atp5d")

for (i in seq_along(gene.idx)) {
  
 p<-DbDynamicsPlot(srt = integrated.xDbit,feature = gene.idx[i],
                   assay = "DeNoise")
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/TCA/",
                            gene.idx[i],".pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/TCA/",
                            gene.idx[i],".png"),
        width = 6,height = 4,dpi = 300)
  
}

# module score
p<-DbDynamicsPlot(srt = integrated.xDbit,
                  feature = "TCA.Module.Score_UCell",
                   assay = "DeNoise")

 ggsave(p,filename = paste0("R figures/scMagSigPro/TCA/TCA module score.pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/TCA/TCA module score.png"),
        width = 6,height = 4,dpi = 300)
 
rm(p)
```

```{r,metabolite,include=FALSE}
load("metabolomics data.RData")


p<-FeatureDimPlot(meta.seu,features = "C6H10O8P",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 0.1,
               lower_quantile =0.05,upper_quantile = 0.9)

ggsave(p,filename = "R figures/scMagSigPro/TCA/TCA metabolite/Glucose-6-phosphate.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/TCA metabolite/Glucose-6-phosphate.pdf",width =15,height = 5,dpi = 300 )


p<-FeatureDimPlot(meta.seu,features = "C21H36O6P",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size =1,
               lower_quantile =0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/TCA/TCA metabolite/mitochondrial membrane phospholipid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/TCA metabolite/mitochondrial membrane phospholipid.pdf",width =15,height = 5,dpi = 300 )

p<-FeatureDimPlot(meta.seu,features = "C10H13N5O7P",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 1,
               lower_quantile = 0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/TCA/TCA metabolite/AMP.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/TCA metabolite/AMP.pdf",width =15,height = 5,dpi = 300 )
###
p<-FeatureDimPlot(meta.seu,features = "C16H31O2",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 1,
               lower_quantile = 0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Palmitic Acid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Palmitic Acid.pdf",width =15,height = 5,dpi = 300 )

p<-FeatureDimPlot(meta.seu,features = "C18H31O2",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 1,
               lower_quantile = 0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Linoleic Acid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Linoleic Acid.pdf",width =15,height = 5,dpi = 300 )

p<-FeatureDimPlot(meta.seu,features = "C18H35O2",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 1,
               lower_quantile = 0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Stearic Acid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Stearic Acid.pdf",width =15,height = 5,dpi = 300 )

p<-FeatureDimPlot(meta.seu,features = "C20H31O2",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 1,
               lower_quantile = 0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Eicosadienoic Acid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/TCA/b-oxidation/Eicosadienoic Acid.pdf",width =15,height = 5,dpi = 300 )


##
meta.seu$age<-factor(meta.seu$age,levels = c("young","old"))

p<-FeatureStatPlot(meta.seu,group.by = "age",palette ="lancet",
                stat.by = c("C6H10O8P",
                            "C21H36O6P",
                            "C10H13N5O7P"),
                assay = "Metabolite",slot = "counts",
                add_box = T,stack = T
                )+
  ggtitle("Impaired TCA Cycle")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/TCA/TCA metabolite/TCA metabolite.pdf",
       height = 6,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/TCA/TCA metabolite/TCA metabolite.png",
       height = 6,width = 6,dpi = 300)

## FeatureStatPlot to check fatty acid accumulation
meta.seu$age<-factor(meta.seu$age,levels = c("young","old"))

p<-FeatureStatPlot(meta.seu,group.by = "age",palette ="lancet",
                stat.by = c("C16H31O2","C18H31O2","C18H35O2","C20H31O2"),
                assay = "Metabolite",slot = "counts",
                add_box = T,stack = T
                )+
  ggtitle("impaired mitochondrial β-oxidation")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/TCA/b-oxidation/b-oxidation.pdf",
       height = 6,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/TCA/b-oxidation/b-oxidation.png",
       height = 6,width = 6,dpi = 300)
```

## bile acids, fatty acid and peroxisomal lipid metabolism
```{r}
# Extract bile acids, fatty acid and peroxisomal lipid metabolism associated gene and calcuate the module score
load("heatmap obj for scMagSigPro.RData")

enrich.res<-ht$enrichment$enrichment

enrich.res<-enrich.res[which(enrich.res$Database=="Reactome" &
                               enrich.res$p.adjust<0.05),]

bile.res<-enrich.res[which(enrich.res$Description=="Bile acid and bile salt metabolism"),]

fatty.res<-enrich.res[which(enrich.res$Description=="Fatty acid metabolism"),]

peroxisomal.res<-enrich.res[which(enrich.res$Description=="Peroxisomal lipid metabolism"),]

bile.gene<-str_split(bile.res$geneID,"/")[[1]]
fatty.gene<-c(str_split(fatty.res$geneID,"/")[[1]],
              str_split(fatty.res$geneID,"/")[[2]])
peroxisomal.gene<-str_split(peroxisomal.res$geneID,"/")[[1]]

gene.list<-list("Bile.Acid.Metabolism.Module.Score"=bile.gene,
                "Fatty.acid.metabolism.Module.Score"=fatty.gene,
                "Peroxisomal.lipid.metabolism.Module.Score"=peroxisomal.gene)

gene.vec<-c(bile.gene,fatty.gene,peroxisomal.gene)
names(gene.vec)<-c(rep("Bile Acid Metabolism",5),
                   rep("Fatty acid Metabolism",43),
                   rep("Peroxisomal Lipid Metabolism",4))

integrated.xDbit<-AddModuleScore_UCell(integrated.xDbit,
                                       assay = "RNA",
                                       features =gene.list,
                                       BPPARAM = BiocParallel::SnowParam(workers = 1),name = "")

# dynamic heatmap for all associated gene

integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =gene.vec,
                                       assay = "DeNoise",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))

gene.vec<-gene.vec[-c(41,44,46,47)]

gene.cluster<-names(gene.vec)
names(gene.cluster)<-gene.vec

ht1 <- DynamicHeatmap(
  srt =integrated.xDbit, features = gene.vec,
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = 25, 
 feature_split =gene.cluster,
  cluster_rows = TRUE,
  min_expcells = 1,
  nlabel = 15,
  show_row_names = F,
  r.sq = 0,
  dev.expl = 0,
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  height = 4, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/metabolism/dynamics heatmap for metabolism.pdf",width = 12,height =6)
print(ht1$plot)
dev.off()

```

```{r,lm for ploidy ,include=FALSE}
source("ScatterHistoPlot.R")
p<-ScatterHistoPlot(seu = integrated.xDbit,
                 var = "Fatty.Acid.Metabolism.Module.Score",
                 plot.x = "nuclei.2n",
                 xlab = "2n Probability",
                 plot.y = "Fatty.Acid.Metabolism.Module.Score",
                 ylab = "Fatty Acid Metabolism Module Score")

ggsave(p,filename = "R figures/scMagSigPro/metabolism/lm between 2n and Fatty Acid Metabolism Module Score.pdf",height = 4,width = 6)
ggsave(p,filename = "R figures/scMagSigPro/metabolism/lm between 2n and Fatty Acid Metabolism Module Score.png",height = 4,width = 6)

p<-ScatterHistoPlot(seu = integrated.xDbit,
                 var = "Fatty.Acid.Metabolism.Module.Score",
                 plot.x = "nuclei.4n",
                 xlab = "4n Probability",
                 plot.y = "Fatty.Acid.Metabolism.Module.Score",
                 ylab = "Fatty Acid Metabolism Module Score")

ggsave(p,filename = "R figures/scMagSigPro/metabolism/lm between 4n and Fatty Acid Metabolism Module Score.pdf",height = 4,width = 6)
ggsave(p,filename = "R figures/scMagSigPro/metabolism/lm between 4n and Fatty Acid Metabolism Module Score.png",height = 4,width = 6)
```

```{r,metablism asso genes for cluster 3,include=FALSE}
integrated.xDbit$age<-factor(integrated.xDbit$age,
                             levels = c("young","old"))

p<-FeatureStatPlot(integrated.xDbit,group.by = "liver.zonation",
                split.by = "age",palette ="lancet",
                stat.by = c("Cyp7a1",
                            "Hsd3b7",
                            "Acaca",
                            "Cyp4a12b",
                            "Elovl5",
                            "Hao2"),
                assay = "DeNoise",
                add_box = T,stack = T
                )+
  ggtitle("Dysfunctional Fatty Acid Metabolism")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/metabolism/metabolism gene.pdf",
       height = 10,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/metabolism/metabolism gene.png",
       height = 10,width = 6,dpi = 300)


source("DbDynamicsPlot.R") 

gene.idx<-c("Cyp7a1", "Hsd3b7", "Acaca","Cyp4a12b", "Elovl5", "Hao2")

for (i in seq_along(gene.idx)) {
  
 p<-DbDynamicsPlot(srt = integrated.xDbit,feature = gene.idx[i],
                   assay = "DeNoise")
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/metabolism/",
                            gene.idx[i],".pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/metabolism/",
                            gene.idx[i],".png"),
        width = 6,height = 4,dpi = 300)
  
}

# module score
p<-DbDynamicsPlot(srt = integrated.xDbit,
                  feature = "Fatty.Acid.Metabolism.Module.Score",
                   assay = "DeNoise")

 ggsave(p,filename = paste0("R figures/scMagSigPro/metabolism/Fatty Acid Metabolism Module Score.pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/metabolism/Fatty Acid Metabolism Module Score.png"),
        width = 6,height = 4,dpi = 300)
 
rm(p)
```

```{r,subset example for vector field,include=FALSE}
Idents(integrated.xDbit)<-"id"
example.xdbit<-subset(integrated.xDbit,idents=c("37_43_C1","37_43_A2"))
save(example.xdbit,file = "example xdbit.RData")
```

```{r,prepare for vf,include=FALSE}
sp.list<-SplitObject(example.xdbit,split.by = "age")

source("PrepareForVF_RNA.R")
sp.list[[1]]<-PrepareForVF_RNA(srt =sp.list[[1]])
sp.list[[2]] <-PrepareForVF_RNA(srt =sp.list[[2]]  )
names(sp.list)<-c("Old","Young")
```


```{r,vector field,include=FALSE}
source("runVectorFields.R")
source("plotVectorFields.R")
#Cyp4a12b
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "Cyp4a12b",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "Cyp4a12b",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/metabolism/Cyp4a12b-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/metabolism/Cyp4a12b-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}

#Acaca
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "Acaca",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "Acaca",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/metabolism/Acaca-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/metabolism/Acaca-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}

#Elovl5
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "Elovl5",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "Elovl5",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/metabolism/Elovl5-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/metabolism/Elovl5-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}
```

```{r,metabolite,include=FALSE}
load("metabolomics data.RData")


p<-FeatureDimPlot(meta.seu,features = "C27H45O4S",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 0.1,
               lower_quantile =0.05,upper_quantile = 0.9)

ggsave(p,filename = "R figures/scMagSigPro/metabolism/fatty acid metabolite/Cholesterol sulfate.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/metabolism/fatty acid metabolite/Cholesterol sulfate.pdf",width =15,height = 5,dpi = 300 )


p<-FeatureDimPlot(meta.seu,features = "C26H44NO7S",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size =1,
               lower_quantile =0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/metabolism/fatty acid metabolite/taurocholic acid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/metabolism/fatty acid metabolite/taurocholic acid.pdf",width =15,height = 5,dpi = 300 )



## FeatureStatPlot to check fatty acid accumulation
meta.seu$age<-factor(meta.seu$age,levels = c("young","old"))

p<-FeatureStatPlot(meta.seu,group.by = "age",palette ="lancet",
                stat.by = c("C27H45O4S","C26H44NO7S"),
                assay = "Metabolite",slot = "counts",
                add_box = T,stack = T
                )+
  ggtitle("Dysfunctional Fatty Acid Metabolism")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/metabolism/fatty acid metabolite/b-oxidation.pdf",
       height = 6,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/metabolism/fatty acid metabolite/b-oxidation.png",
       height = 6,width = 6,dpi = 300)

```


## immune response
```{r,Extract immune response and calcuate the module score,include=FALSE}
# Extract immune response and calcuate the module score
load("heatmap obj for scMagSigPro.RData")

enrich.res<-ht$enrichment$enrichment

enrich.res<-enrich.res[which(enrich.res$Database=="Reactome" &
                               enrich.res$p.adjust<0.05),]

Hemostasis.res<-enrich.res[which(enrich.res$Description=="Hemostasis"),]
Platelet.res<-enrich.res[which(enrich.res$Description=="Platelet activation, signaling and aggregation"),]
DAP12.res<-enrich.res[which(enrich.res$Description=="DAP12 interactions"),]
Immunoregulatory.res<-enrich.res[which(enrich.res$Description=="Immunoregulatory interactions between a Lymphoid and a non-Lymphoid cell"),]

Hemostasis.gene<-str_split(Hemostasis.res$geneID,"/")[[1]]
Platelet.gene<-c(str_split(Platelet.res$geneID,"/")[[1]])
DAP12.gene<-str_split(DAP12.res$geneID,"/")[[1]]
Immunoregulatory.gene<-str_split(Immunoregulatory.res$geneID,"/")[[1]]

gene.list<-list("Hemostasis.Module.Score"=Hemostasis.gene,
                "Platelet.Activation.Module.Score"=Platelet.gene,
                "DAP12.Interactions.Module.Score"=DAP12.gene,
                "Immunoregulatory.Interactions.Module.Score"=Immunoregulatory.gene)



integrated.xDbit<-AddModuleScore_UCell(integrated.xDbit,
                                       assay = "RNA",
                                       features =gene.list,
                                       BPPARAM = BiocParallel::SnowParam(workers = 1),name = "")
```

```{r,DynamicHeatmap,include=FALSE}

integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =Hemostasis.gene,
                                       assay = "DeNoise",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))

DefaultAssay(integrated.xDbit)<-"DeNoise"
ht1 <- DynamicHeatmap(
  srt =integrated.xDbit, features = Hemostasis.gene,
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = 25, 
  cluster_rows = TRUE,
  min_expcells = 1,
  nlabel = 10,
  show_row_names = F,
  r.sq = 0,
  dev.expl = 0,
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  height =4, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/immune/dynamics heatmap for metabolism(Hemostasis).pdf",width =8,height = 6)
print(ht1$plot)
dev.off()


integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =Platelet.gene,
                                       assay = "DeNoise",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))

ht1 <- DynamicHeatmap(
  srt =integrated.xDbit, features = Platelet.gene,
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = 25, 
  cluster_rows = TRUE,
  min_expcells = 1,
  nlabel = 10,
  show_row_names = F,
  r.sq = 0,
  dev.expl = 0,
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  height = 4, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/immune/dynamics heatmap for metabolism(Platelet).pdf",width =8,height = 6)
print(ht1$plot)
dev.off()

integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =DAP12.gene,
                                       assay = "DeNoise",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))

ht1 <- DynamicHeatmap(
  srt =integrated.xDbit, features = DAP12.gene,
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = 25, 
  cluster_rows = TRUE,
  min_expcells = 1,
  nlabel = 10,
  show_row_names = F,
  r.sq = 0,
  dev.expl = 0,
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  height = 4, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/immune/dynamics heatmap for metabolism(DAP12).pdf",width =8,height =6)
print(ht1$plot)
dev.off()

```

```{r,immune asspciated gene ,include=FALSE}
source("DbDynamicsPlot.R") 

gene.idx<-c("Gata6", "Itgb3", "Clu","Pecam1", "Clec1b", "Fcer1g","Fgg","Fga",
           "Sirpa", "H2-Q7", "H2-Q6","Tyrobp")

for (i in seq_along(gene.idx)) {
  
 p<-DbDynamicsPlot(srt = integrated.xDbit,feature = gene.idx[i],
                   assay = "DeNoise")
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/",
                            gene.idx[i],".pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/",
                            gene.idx[i],".png"),
        width = 6,height = 4,dpi = 300)
  
}

# module score
p<-DbDynamicsPlot(srt = integrated.xDbit,
                  feature =  "DAP12.Interactions.Module.Score",
                   assay = "DeNoise")

 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/DAP12 Interactions Module Score.pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/DAP12 Interactions Module Score.png"),
        width = 6,height = 4,dpi = 300)
 
 p<-DbDynamicsPlot(srt = integrated.xDbit,
                  feature =  "Hemostasis.Module.Score",
                   assay = "DeNoise")

 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/Hemostasis Module Score.pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/Hemostasis Module Score.png"),
        width = 6,height = 4,dpi = 300)
 
 p<-DbDynamicsPlot(srt = integrated.xDbit,
                  feature =  "Platelet.Activation.Module.Score",
                   assay = "DeNoise")

 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/Platelet Activation Module Score.pdf"),
        width = 6,height = 4,dpi = 300)
 
 ggsave(p,filename = paste0("R figures/scMagSigPro/immune/Platelet Activation Module Score.png"),
        width = 6,height = 4,dpi = 300)
 
 
 
rm(p)
## featurestatplot
integrated.xDbit$age<-factor(integrated.xDbit$age,
                             levels = c("young","old"))

p<-FeatureStatPlot(integrated.xDbit,group.by = "liver.zonation",
                split.by = "age",palette ="lancet",
                stat.by = c("Gata6", "Itgb3", "Clu",
                            "Pecam1", "Clec1b", 
                            "Fcer1g","Fgg","Fga","Sirpa", 
                            "H2-Q7", "H2-Q6","Tyrobp"),
                assay = "DeNoise",
                add_box = T,stack = T
                )+
  ggtitle("Dysfunctional immune response")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/immune/immune gene.pdf",
       height = 10,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/immune/immune gene.png",
       height = 10,width = 6,dpi = 300)

```

```{r,prepare for vf,include=FALSE}
load("example xdbit.RData")
DefaultAssay(example.xdbit)<-"DeNoise"
sp.list<-SplitObject(example.xdbit,split.by = "age")

source("PrepareForVF_RNA.R")
sp.list[[1]]<-PrepareForVF_RNA(srt =sp.list[[1]])
sp.list[[2]] <-PrepareForVF_RNA(srt =sp.list[[2]]  )
names(sp.list)<-c("Old","Young")
```

```{r,vector field,include=FALSE}
source("runVectorFields.R")
source("plotVectorFields.R")
#Gata6
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "Gata6",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "Gata6",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/immune/Gata6-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/immune/Gata6-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}

#Fgg
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "Fgg",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "Fgg",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/immune/Fgg-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/immune/Fgg-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}

#Fga
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "Fga",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "Fga",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/immune/Fga-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/immune/Fga-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}

#H2-Q7
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "H2-Q7",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "H2-Q7",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/immune/H2-Q7-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/immune/H2-Q7-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}

#H2-Q6
for (i in 1:2) {
  
  vf<-runVectorFields(sp.list[[i]],features = "H2-Q6",dist.spot=5)

  p<-plotVectorFields(VF = vf,parameter = "H2-Q6",pt.size = 3)
  
  ggsave(p,filename =paste0("R figures/scMagSigPro/immune/H2-Q6-VF-",
                            names(sp.list)[i],".pdf"),dpi = 150,
         width = 6,height = 5)
    ggsave(p,filename =paste0("R figures/scMagSigPro/immune/H2-Q6-VF-",
                            names(sp.list)[i],".png"),dpi = 150,
         width = 6,height = 5)
  
}
```

```{r,metabolite,include=FALSE}
load("metabolomics data.RData")


p<-FeatureDimPlot(meta.seu,features = "C18H29O3S",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size = 0.1,
               lower_quantile =0.05,upper_quantile = 0.9)

ggsave(p,filename = "R figures/scMagSigPro/immune/Arachidonic acid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/immune/Arachidonic acid.pdf",width =15,height = 5,dpi = 300 )


p<-FeatureDimPlot(meta.seu,features = "C19H15O4",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size =1,
               lower_quantile =0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/immune/Flavonoid.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/immune/Flavonoid.pdf",width =15,height = 5,dpi = 300 )


p<-FeatureDimPlot(meta.seu,features = "C23H47NO7P",
                  reduction = "spatial",palette = "RdYlBu",
               slot = "counts",aspect.ratio =0.3,pt.size =1,
               lower_quantile =0.05,upper_quantile = 0.95)

ggsave(p,filename = "R figures/scMagSigPro/immune/Lysophosphatidylcholine.png",width =15,height = 5,dpi = 300 )

ggsave(p,filename = "R figures/scMagSigPro/immune/Lysophosphatidylcholine.pdf",width =15,height = 5,dpi = 300 )


## FeatureStatPlot to check fatty acid accumulation
meta.seu$age<-factor(meta.seu$age,levels = c("young","old"))

p<-FeatureStatPlot(meta.seu,group.by = "age",palette ="lancet",
                stat.by = c("C18H29O3S","C19H15O4","C23H47NO7P"),
                assay = "Metabolite",slot = "counts",
                add_box = T,stack = T
                )+
  ggtitle("Dysfunctional Immune Response")+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,size = 14)
        )

ggsave(p,filename ="R figures/scMagSigPro/immune/Dysfunctional Immune Response.pdf",
       height = 6,width = 6,dpi = 300)
ggsave(p,filename ="R figures/scMagSigPro/immune/Dysfunctional Immune Response.png",
       height = 6,width = 6,dpi = 300)

```

```{r,cell prob,include=FALSE}
cell.prob<-t(integrated.xDbit@meta.data[,c(16:25)])
cell.prob<-CreateAssayObject(data =cell.prob,min.cells = 0,min.features = 0 )
integrated.xDbit[["cell.prob"]]<-cell.prob

integrated.xDbit$age<-factor(integrated.xDbit$age,levels = c("young","old"))

idx<-rownames(integrated.xDbit@assays$cell.prob)

for (i in seq_along(idx)) {

p<-FeatureStatPlot(integrated.xDbit,assay = "cell.prob",
                slot = "data",stat.by = idx[i],add_box = T,
                group.by = "liver.zonation",split.by = "age",
                comparisons = list(c("young","old")),title = "",
                multiplegroup_comparisons = T,palette = "lancet")+
  ylab("Cell Probablity")+xlab("")+

theme(axis.title.y = element_text(hjust = 0.5,
                                  size = 14))

ggsave(filename = paste0("R figures/scMagSigPro/immune/",idx[i],
       " probablity.pdf"),width = 5,height = 4,dpi = 300)

ggsave(filename = paste0("R figures/scMagSigPro/immune/",idx[i],
       " probablity.png"),width = 5,height = 4,dpi = 300)
  
}
```

```{r,validation,include=FALSE}
#load("H:/Celia group/xDbit(s4) project(renew)/paired scRNA.RData")

CellDimPlot(pair.sc,reduction = "umap",group.by = "nuclei")

pair.sc$Age<-factor(pair.sc$Age,levels = c("Young","Old"))
pair.sc$nuclei<-factor(pair.sc$nuclei)

p<-CellStatPlot(srt =pair.sc, stat.by = "nuclei", 
                title = "Nuclei Proportion in Peter Tessarz",
             group.by = "Age", label = TRUE,palette = "nejm")

ggsave(p,filename = "R figures/scMagSigPro/immune/Nuclei Proportion in Peter Tessarz.pdf",width = 6,height = 5,dpi = 300)
ggsave(p,filename = "R figures/scMagSigPro/immune/Nuclei Proportion in Peter Tessarz.png",width = 6,height = 5,dpi = 300)

TMS.scRNA<-readRDS(file = "TMS scRNA.rds")

p<-CellStatPlot(srt =TMS.scRNA, stat.by = "cell_type",
                title = "Cell Proportion in TMS",
             group.by = "age", label = TRUE,palette = "jama")

ggsave(p,filename = "R figures/scMagSigPro/immune/Cell Proportion in TMS.pdf",width = 6,height = 5,dpi = 300)

ggsave(p,filename = "R figures/scMagSigPro/immune/Cell Proportion in TMS.png",width = 6,height = 5,dpi = 300)

```





# Pathway Enrichment for scMagSigPro
```{r}
load("heatmap obj for scMagSigPro.RData")

enrichment.res<-as.data.frame(ht[["enrichment"]][["enrichment"]])

write.csv(enrichment.res,file = "R figures/scMagSigPro/GO_BP enrichment.csv")
```

```{r,plot,include=FALSE}
enrichment.res<-ht[["enrichment"]]

idx<-c(paste0("Cluster_",seq(1,6,1)))

for (i in seq_along(idx)) {
  pdf(file =paste0("R figures/scMagSigPro/GO_BP-",idx[i],".pdf"),
      width = 15,height = 7.5)
 print( EnrichmentPlot(res = enrichment.res,db = "GO_BP",
                 plot_type = "network",
                 group_use = idx[i])
 )
  
  dev.off()
  
}


for (i in seq_along(idx)) {
  pdf(file =paste0("R figures/scMagSigPro/Reactome-",
                   idx[i],".pdf"),
      width = 15,height = 7.5)
 print( EnrichmentPlot(res = enrichment.res,db = "Reactome",
                 plot_type = "network",
                 group_use = idx[i])
 )
  
  dev.off()
  
}
```



# scMagSigPro at Rectome

```{r,include=FALSE}
DefaultAssay(integrated.xDbit)<-"Metabolism"
xdbit.sce<-as.SingleCellExperiment(integrated.xDbit)
```

```{r,include=FALSE}
xdbit.meta.mag.obj<-as_scmp(object = xdbit.sce,from = "sce",
                       align_pseudotime = T,
                       verbose = T,
                       additional_params = list(labels_exist = TRUE,
                                                exist_ptime_col = "PV.CV.Trajectory",
                                                exist_path_col = "age"))

xdbit.meta.mag.obj <- sc.squeeze(xdbit.meta.mag.obj)

rm(xdbit.sce)
```

```{r,plot bin,include=FALSE}
p<-plotBinTile(xdbit.meta.mag.obj)+
  ggtitle("Binned Pseudotime by scMagSigPro")+
  ylab("Group")+xlab("Binned Pseudotime ")+
  scale_fill_viridis(name = "Bin Size")+
  theme(plot.title = element_text(face = "bold",size = 14,hjust = 0.5),
        axis.title =element_text(face = "bold",size = 12),
        axis.text = element_text(size = 12))
ggsave(p,filename = "R figures/scMagSigPro/pseudotime bins for metabolic.pdf",width = 6,height = 4,dpi = 300)
ggsave(p,filename = "R figures/scMagSigPro/pseudotime bins for metabolic.png",width = 6,height = 4,dpi = 300)  
```

```{r,set degree,include=FALSE}
xdbit.meta.mag.obj<- sc.set.poly(xdbit.meta.mag.obj, poly_degree =3)

save(xdbit.meta.mag.obj,file ="scMagSigPro obj(metabolic).RData" )
```

```{r,run scMagSigPro,include=FALSE}
# Detect non-flat profiles
xdbit.meta.mag.obj <- sc.p.vector(xdbit.meta.mag.obj,
                             offset = TRUE, 
                             p_value = 0.05, 
                             min_na = 10,
                             verbose =T,
                             log_offset = TRUE
)

# Model refinement

xdbit.meta.mag.obj <- sc.t.fit(xdbit.meta.mag.obj, verbose = TRUE)

save(xdbit.meta.mag.obj,file ="scMagSigPro obj(metabolic).RData" )
```

```{r,get sig gene,include=FALSE}
xdbit.meta.mag.obj  <- sc.filter(
  scmpObj = xdbit.meta.mag.obj ,
  rsq = 0.7,
  vars = "groups",
  intercept = "dummy",
  includeInflu = F
)
```


```{r,cluster the sig scMagSigPro gene,include=FALSE}
xdbit.meta.mag.obj <- sc.cluster.trend(
 xdbit.meta.mag.obj ,
  cluster_by = "counts",
  geneSet = "union",
  k =6
)

p<-plotTrendCluster(
  scmpObj =xdbit.meta.mag.obj ,
  plot = "counts",
  logs = TRUE,
  summary_mode='mean',
  smoothness = 0.01,
  verbose = FALSE,
  lines = TRUE
)

p<-p+
  ggtitle("Dynamical Reactome Pathway over PV-CV axis")+ylab("Binned Pseudobulk Expression")+
  xlab("Pooled PV-CV Trajectory")+
  theme(plot.title = element_text(face = "bold",size = 14,hjust = 0.5),
        axis.text.x = element_text(angle = 0),
        axis.title =element_text(face = "bold",size = 12),
        axis.text = element_text(size = 12))


ggsave(p,filename = "R figures/scMagSigPro/clustered scMagSigPro Sig Reactome pathway.png",width = 7,height = 5,dpi = 300) 
ggsave(p,filename = "R figures/scMagSigPro/clustered scMagSigPro Sig Reactome pathway.pdf",width = 7,height = 5,dpi = 300) 
```

```{r,extract the sig genes}
sig.mag.pathway<-as.data.frame( do.call(rbind,xdbit.meta.mag.obj@Significant@clusters))

sig.mag.pathway$pathway=rownames(sig.mag.pathway)
colnames(sig.mag.pathway)[1]<-"cluster"

save(sig.mag.pathway,file = "sig scMagSigPro pathway.RData")
write.csv(sig.mag.pathway,file = "sig scMagSigPro pathway.csv")
```

```{r,include=FALSE}
integrated.xDbit <- RunDynamicFeatures(srt = integrated.xDbit, 
                                       lineages = c("Young.Trajectory",                                                   "Old.Trajectory"),
                                       features =sig.mag.pathway$pathway,assay = "Metabolism",
                                       BPPARAM = BiocParallel::SnowParam(workers = 1))
```

```{r}
cluster.gene<-paste0("Cluster_",sig.mag.pathway$cluster)
names(cluster.gene)<-sig.mag.pathway$pathway

ht <- DynamicHeatmap(
  srt =integrated.xDbit, 
  lineages = c("Young.Trajectory", "Old.Trajectory"),
  use_fitted = TRUE,  reverse_ht = F,
  pseudotime_label = F, 
  nlabel = 0,
  show_row_names = T,
  feature_split = cluster.gene,
  cluster_rows = TRUE,
  min_expcells = 1,
  r.sq = 0,
  dev.expl = 0,
  heatmap_palette = "turbo",
  padjust = 1,pseudotime_palette = "plasma",
  cell_annotation =c("age","liver.zonation"),
  cell_annotation_palette = c("lancet","npg"),
  topTerm =10,
  label_size = 4,
  height = 10, width = 2,
  use_raster = T,
  raster_by_magick = T,
)
pdf(file = "R figures/scMagSigPro/dynamics metabolic heatmap.pdf",width =16,height = 12)
print(ht$plot)
dev.off()

save(integrated.xDbit,file = "filter data.Rdata")

save(ht,file="heatmap obj for metabolic scMagSigPro.RData")
```

# Cell-type Decovolution(RCTD)

## prepare the ref

```{r,prepare the ref}
dir.list<- "H:/Celia group/xDbit project/ref_data/rawData_mouseStSt/countTable_mouseStSt/"
 
matrix <- as.matrix(readMM(file = paste0(dir.list,"matrix.mtx")))
features <- as.vector(read.table(paste0(dir.list,"features.tsv")))
barcodes <- as.vector(read.table(paste0(dir.list,"barcodes.tsv")))

colnames(matrix)<-barcodes$V1
rownames(matrix)<-features$V1

ref.data<-CreateSeuratObject(matrix ,min.cells = 0,min.features = 0,project = "Ref") 
ref.data$library<-"Ref"
 
 
meta.dat<-read.csv(file = "H:/Celia group/xDbit project/ref_data/annot_mouseStStAll.csv",header = T )

rownames(meta.dat)<-meta.dat$cell

rm(matrix)

ref.data<-AddMetaData(ref.data,metadata = meta.dat)

umap.df<-as.data.frame(meta.dat[,c(1:2)])
rownames(umap.df)<-meta.dat$cell

ref.data[["umap"]]<- CreateDimReducObject(embeddings = as.matrix(umap.df), 
                                           key = "UMAP_", 
                                           assay = "RNA")

save(ref.data,file = "scRNA ref data.RData")

```


```{r,check the ref,include=FALSE}
ref.data$celltype<-ref.data$annot

CellDimPlot(ref.data,group.by = "celltype",reduction = "umap",
            label = T,
            palette = "Set1")+
  ggtitle("Reference(scRNA)")+
      theme(plot.title = element_text( hjust = 0.7,face = "bold",size = 20))
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/RCTD/Dimplot for ref scRNA.pdf",dpi = 150,width = 8,height = 6)
ggsave(filename = "H:/Celia group/xDbit(s4) project(renew)/R figures/RCTD/Dimplot for ref scRNA.png",dpi = 150,width = 8,height = 6)
```

## re-annotation

```{r}
Idents(ref.data)<-"annot"
ref.data<- RenameIdents(ref.data, 
                        'pDCs' = 'Immune',
                        'T cells' = 'Immune', 
                        'Monocytes & Monocyte-derived cells' = 'Immune',
                        'ILC1s' = 'Immune', 
                        'B cells' = 'Immune',
                        'cDC2s' = 'Immune', 
                        'NK cells' = 'Immune',
                        'Neutrophils' = 'Immune', 
                        'Mig. cDCs' = 'Immune',
                        'cDC1s' = 'Immune', 
                        'Basophils' = 'Immune')
ref.data$main.celltype<-ref.data@active.ident

Idents(ref.data)<-"main.celltype"

hep<-subset(ref.data,idents="Hepatocytes")


VlnPlot(hep,features =c("Cyp2f2","Cyp2e1","Glul","Hamp","Igfbp2","Adh1b","Adh4","Gsta1","Adh1c"),group.by ="sub.cluster",stack = T,sort = T,flip = T)


Idents(ref.data)<-"sub.cluster"

ref.data<-RenameIdents(ref.data, 
                        "Hepatocytes_13" = "Hepatocytes_0",
                        "Hepatocytes_8" = "Hepatocytes_0", 
                        "Hepatocytes_14" = "Hepatocytes_1",
                        "Hepatocytes_5"= "Hepatocytes_2", 
                        "Hepatocytes_10" = "Hepatocytes_2",
                        "Hepatocytes_9" = "Hepatocytes_2", 
                       "Hepatocytes_11" = "Hepatocytes_2",
                        "Hepatocytes_12" = "Hepatocytes_2", 
                        "Hepatocytes_6" = "Hepatocytes_3",
                        "Hepatocytes_7" = "Hepatocytes_3", 
                        "Hepatocytes_15" = "Hepatocytes_4" )
ref.data$sub.cluster<-ref.data@active.ident



Idents(hep)<-"sub.cluster"

hep<-RenameIdents(hep, 
                        "Hepatocytes_13" = "Hepatocytes_0",
                        "Hepatocytes_8" = "Hepatocytes_0", 
                        "Hepatocytes_14" = "Hepatocytes_1",
                        "Hepatocytes_5"= "Hepatocytes_2", 
                        "Hepatocytes_10" = "Hepatocytes_2",
                        "Hepatocytes_9" = "Hepatocytes_2", 
                       "Hepatocytes_11" = "Hepatocytes_2",
                        "Hepatocytes_12" = "Hepatocytes_2", 
                        "Hepatocytes_6" = "Hepatocytes_3",
                        "Hepatocytes_7" = "Hepatocytes_3", 
                        "Hepatocytes_15" = "Hepatocytes_4" )
hep$sub.cluster<-hep@active.ident


VlnPlot(hep,features =c("Cyp2f2","Acly","Asl","Uox",
                        "Hamp","Igfbp2",
                        "Cyp2e1","Glul","Cyp1a2","Cyp27a1",
                        "Gsta3","Ugt1a1"),
        group.by ="sub.cluster",stack = T,sort = T,flip = T)

Idents(ref.data)<-"sub.cluster"

ref.data<-RenameIdents(ref.data, 
                        "Hepatocytes_3" = "Hepatocytes_1",
                        "Hepatocytes_4" = "Hepatocytes_1" 
                        )
ref.data$sub.cluster<-ref.data@active.ident

Idents(hep)<-"sub.cluster"

hep<-RenameIdents(hep, 
                        "Hepatocytes_3" = "Hepatocytes_1",
                        "Hepatocytes_4" = "Hepatocytes_1" 
                        )
hep$sub.cluster<-hep@active.ident

hep$sub.cluster<-as.character(hep$sub.cluster)


p<-CellDimPlot(hep,group.by = "sub.cluster",split.by = "sub.cluster" ,
            reduction = "umap",ncol = 1,
            label = T,raster = F,
            palette = "Set1")
ggsave(p,filename = "tmp3.pdf",height = 20,width = 7,units = "in")
FeatureDimPlot(hep,features ="Cyp2f2" ,reduction = "umap",palette = "Reds")
FeatureDimPlot(hep,features ="Hamp" ,reduction = "umap",palette = "Reds")
FeatureDimPlot(hep,features ="Cyp2e1",reduction = "umap",palette = "Reds")


VlnPlot(hep,features =c("Cyp2f2","Acly","Asl","Uox",
                        "Hamp","Igfbp2",
                        "Cyp2e1","Glul","Cyp1a2","Cyp27a1",
               "Gsta3","Ugt1a1"),group.by ="sub.cluster",stack = T,sort = T,flip = T)
ggsave("tmp5.pdf",width = 5,height = 12)


hep<-RunDM(hep,reduction = "pca",dims = 1:10)

CellDimPlot(hep,group.by = "sub.cluster",reduction = "dm",
            label = T,
            palette = "Set1")
ggsave(filename = "tmp6.pdf",width = 8)


Idents(ref.data)<-"sub.cluster"

ref.data<-RenameIdents(ref.data, 
                        "Hepatocytes_0" = "Hepatocytes.CV",
                        "Hepatocytes_1" = "Hepatocytes.PV",
                        "Hepatocytes_2" = "Hepatocytes.CV" 
                        )
ref.data$sub.cluster<-ref.data@active.ident

Idents(hep)<-"sub.cluster"

hep<-RenameIdents(hep, 
                        "Hepatocytes_0" = "Hepatocytes.CV",
                        "Hepatocytes_1" = "Hepatocytes.PV",
                        "Hepatocytes_2" = "Hepatocytes.CV" 
                        )
hep$sub.cluster<-hep@active.ident

hep$sub.cluster<-as.character(hep$sub.cluster)

ref.data$main.celltype<-as.character(ref.data$sub.cluster)
hep$main.celltype<-as.character(hep$sub.cluster)

CellDimPlot(ref.data,group.by = "main.celltype",reduction = "umap",
            label = T,
            palette = "Set1")+
  ggtitle("Reference(revised annotation)")+
      theme(plot.title = element_text( hjust = 0.7,face = "bold",size = 20))
ggsave(filename = "H:/Celia group/xDbit project/R figures/RCTD/Dimplot for ref scRNA(revised).pdf",dpi = 150,width = 8,height = 6)
ggsave(filename = "H:/Celia group/xDbit project/R figures/RCTD/Dimplot for ref scRNA(revised).png",dpi = 150,width = 8,height = 6)


CellDimPlot(hep,group.by = "main.celltype",reduction = "umap",
            label = T,theme_use = "theme_blank",
            palette = "Set1")
ggsave(filename = "H:/Celia group/xDbit project/R figures/RCTD/Dimplot for ref scRNA(revised)-Hepatocyte.pdf",dpi = 150,width = 8,height = 6)
ggsave(filename = "H:/Celia group/xDbit project/R figures/RCTD/Dimplot for ref scRNA(revised)-Hepatocyte.png",dpi = 150,width = 8,height = 6)

feature<-c(c("Cyp2f2","Acly","Asl","Uox",
              "Hamp","Igfbp2",
              "Cyp2e1","Glul","Cyp1a2","Cyp27a1","Gsta3"))

VlnPlot(hep,features =feature,group.by ="main.celltype",stack = T,
        sort = F,flip = T)
ggsave("H:/Celia group/xDbit project/R figures/RCTD/Vlnplot for ref scRNA(revised)-Hepatocyte.png",width = 6,height = 12)

ggsave("H:/Celia group/xDbit project/R figures/RCTD/Vlnplot for ref scRNA(revised)-Hepatocyte.pdf",width = 6,height = 12)

save(ref.data,file = "scRNA ref data.RData")

```




## prepare the input for RCTD

```{r,include=FALSE}
## query

load("H:/Celia group/xDbit(s4) project(renew)/filter data.RData")

counts<-integrated.xDbit@assays$RNA@counts

coords<-integrated.xDbit@reductions[["spatial"]]@cell.embeddings

colnames(coords) <- c("x", "y")

coords[is.na(colnames(coords))] <- NULL

coords<-as.data.frame(coords)

query<-SpatialRNA(coords = coords,counts = counts,colSums(counts))

rm(coords,counts,integrated.xDbit)

## reference

load("H:/Celia group/xDbit(s4) project(renew)/scRNA ref data.RData")

Idents(ref.data)<-"main.celltype"

counts <- ref.data@assays$RNA@counts

cluster <- as.factor(ref.data$main.celltype)

names(cluster) <- colnames(ref.data)

nUMI <- ref.data$nCount_RNA

names(nUMI) <- colnames(ref.data)

reference <- Reference(counts, cluster, nUMI,n_max_cells =50000)

save(query,reference,file = "RCTD input.RData")

```

## Run RCTD

```{r,run RCTD,include=FALSE}
# full mode
rm(list=ls())

gc()

load("H:/Celia group/xDbit(s4) project(renew)/RCTD input.RData")

RCTD <- create.RCTD(query, reference, max_cores = 8)

rm(query,reference)

gc()

RCTD <- run.RCTD(RCTD, doublet_mode = 'full')

save(RCTD,file = "RCTD output(full).RData")

```
## Check the results and put back to spatial obj

```{r,multi mode,include=FALSE}
load("H:/Celia group/xDbit(s4) project(renew)/RCTD output(full).RData")

barcodes <- colnames(RCTD@spatialRNA@counts)

weights <- RCTD@results$weights

norm_weights <- normalize_weights(weights)

cell_types <-RCTD@cell_type_info[["info"]][[2]]

decov.res<-as.data.frame(norm_weights[,cell_types])

integrated.xDbit<-AddMetaData(integrated.xDbit,metadata = decov.res,
                              col.name = colnames(decov.res))

integrated.xDbit@misc$decov.full.res<-decov.res


save(integrated.xDbit,file = "filter data.Rdata")

write.csv(decov.res,file = "cell-type decov res.csv")

```

```{r,add ct decov assay,include=FALSE}
ct.deconv.assay <-CreateAssayObject(data = as.matrix(t(decov.res)),
                              min.cells = 0,min.features = 0)

integrated.xDbit[["ctDeconv"]]<-ct.deconv.assay

DefaultAssay(integrated.xDbit)<-"ctDeconv"

all.features<-rownames(integrated.xDbit@assays$ctDeconv@data)

integrated.xDbit<-ScaleData(integrated.xDbit,assay = "ctDeconv",
                            features = all.features)

integrated.xDbit<-RunPCA(integrated.xDbit,assay ="ctDeconv",
                         features = all.features,
                         reduction.name ="ctDeconvPCA",
                          reduction.key = "ctDeconvPCA_")

integrated.xDbit<-RunUMAP(integrated.xDbit,reduction ="ctDeconvPCA",
                          dims = 1:5,n.neighbors=100L,min.dist=0.001,
                          reduction.name = "ctDeconvUMAP",
                          reduction.key = "ctDeconvUMAP_")

CellDimPlot(integrated.xDbit,group.by = "age",reduction = "ctDeconvPCA")
CellDimPlot(integrated.xDbit,group.by = "age",reduction = "ctDeconvUMAP")
```
```{r,heatmap for ct deconv,include=FALSE}
pdf(file = "ct deconv.pdf",width = 6,height = 5)
GroupHeatmap(integrated.xDbit,assay ="ctDeconv",exp_method ="zscore",
             slot = "data",features = all.features,group.by = c("liver.zonation"),split.by = c("age"))
dev.off()
```


## Visualization for full mode

```{r}
spatial.list<-SplitObject(integrated.xDbit,split.by = "id")
```

```{r,piechart in spatial,include=FALSE}
source("DbPie.R")

color.vec<-c(brewer.pal(8,"Set3"))

for (i in 1:length(spatial.list)) {
  
  tmp<-spatial.list[[i]]
  
  m <- as.matrix(tmp@meta.data[,c(16:23)])
  p <- tmp@reductions[["spatial"]]@cell.embeddings
  colnames(p)<-c("x","y")
  
  r<-if_else(grepl("37_30",unique(tmp$id)),true = 100,false = 150)
  
  p1<-Db.pie(m,p,color.vec,r = r)+
      ggtitle(unique(tmp$id))+
      theme(plot.background = element_rect(fill = "black", color = "black"),
            panel.background = element_rect(fill = "black", color = "black"),
            plot.margin = unit(c(1,1,1,1), "lines"), 
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            text = element_text(color = "white"),
            axis.title.x = element_blank(), 
            axis.title.y = element_blank(),  
            axis.text.x = element_blank(),  
            axis.text.y = element_blank(),   
            axis.ticks = element_blank() ,
            legend.background = element_rect(fill = "black", color = "black"),
            legend.title = element_text(color = "white",size = 20,face = "bold"), 
            legend.text = element_text(color = "white",size = 18), 
            legend.key.size = unit(3, "cm"),  
            plot.title = element_text(color = "white", hjust = 0.5,
                                      face = "bold",size = 40))+
      theme(aspect.ratio = 1)
  
  

ggsave(p1,filename =paste0("H:/Celia group/xDbit(s4) project(renew)/R figures/RCTD/PieChart-celltype deconv-",unique(tmp$id),".pdf" ),
                            width = 25,height = 25,units = "in",dpi = 150)
ggsave(p1,filename =paste0("H:/Celia group/xDbit(s4) project(renew)/R figures/RCTD/PieChart-celltype deconv-",unique(tmp$id),".png" ),
                            width = 25,height = 25,units = "in",dpi = 150)

rm(tmp,m,p,p1)

}

```


# nuclei Decovolution(RCTD)
## prepare the input for RCTD

```{r,include=FALSE}
## query

load("H:/Celia group/xDbit(s4) project(renew)/filter data.RData")

counts<-integrated.xDbit@assays$RNA@counts

coords<-integrated.xDbit@reductions[["spatial"]]@cell.embeddings

colnames(coords) <- c("x", "y")

coords[is.na(colnames(coords))] <- NULL

coords<-as.data.frame(coords)

query<-SpatialRNA(coords = coords,counts = counts,colSums(counts))

rm(coords,counts,integrated.xDbit)

## reference

load("H:/Celia group/xDbit(s4) project(renew)/paired scRNA.RData")

Idents(pair.sc)<-"nuclei"

pair.sc<-subset(pair.sc,idents=c("2n" ,"4n"))

counts <- pair.sc@assays$RNA@counts

cluster <- as.factor(pair.sc$nuclei)

names(cluster) <- colnames(pair.sc)

nUMI <- pair.sc$nCount_RNA

names(nUMI) <- colnames(pair.sc)

reference <- Reference(counts, cluster, nUMI,n_max_cells =50000)

save(query,reference,file = "RCTD input(nuclei).RData")

```
## Run RCTD
```{r,run RCTD,include=FALSE}
# full mode
rm(list=ls())

gc()

load("RCTD input(nuclei).RData")

RCTD <- create.RCTD(query, reference, max_cores = 8)

rm(query,reference)

gc()

RCTD <- run.RCTD(RCTD, doublet_mode = 'full')

save(RCTD,file = "RCTD nuclei output(full).RData")

```
## Check the results and put back to spatial obj

```{r,multi mode,include=FALSE}
load("RCTD nuclei output(full).RData")

barcodes <- colnames(RCTD@spatialRNA@counts)

weights <- RCTD@results$weights

norm_weights <- normalize_weights(weights)

cell_types <-RCTD@cell_type_info[["info"]][[2]]

nuclei.decov.res<-as.data.frame(norm_weights[,cell_types])

colnames(nuclei.decov.res)<-c("nuclei.2n","nuclei.4n")

integrated.xDbit<-AddMetaData(integrated.xDbit,
                              metadata = nuclei.decov.res,
                              col.name = colnames(nuclei.decov.res))

integrated.xDbit@misc$decov.nuclei.res<-nuclei.decov.res


save(integrated.xDbit,file = "filter data.Rdata")

write.csv(nuclei.decov.res,file = "nuclei decov res.csv")

```

```{r,add nuc decov assay,include=FALSE}
nuc.deconv.assay <-CreateAssayObject(data = as.matrix(t(nuclei.decov.res)),
                              min.cells = 0,min.features = 0)

integrated.xDbit[["nucDeconv"]]<-nuc.deconv.assay

DefaultAssay(integrated.xDbit)<-"nucDeconv"

all.features<-rownames(integrated.xDbit@assays$nucDeconv@data)

integrated.xDbit<-ScaleData(integrated.xDbit,assay = "nucDeconv",
                            features = all.features)

colnames(nuclei.decov.res)<-c("Ploidy_1","Ploidy_2")

rna.embeding <- CreateDimReducObject(
  embeddings =as.matrix(nuclei.decov.res),
  assay = "RNA",
  key = "Ploidy_")

integrated.xDbit[['ploidy']] <- rna.embeding


save(integrated.xDbit,file = "filter data.Rdata")
```

```{r}
FeatureDimPlot(integrated.xDbit,feature = "nuclei.2n",reduction = "umap")
CellDimPlot(integrated.xDbit,group.by = "liver.zonation",reduction = "ploidy",pt.size = 1.2)
```







